---
title: "Project_3_Production"
author: "Bea Sauve, Elton nichols, Kainen Osborne"
date: "2026-03-03"
output: html_document
---

# Project 3 Production Notebook
This notebook will entail all team members work including data refinement, forecasting productions, and statistic generations. Team member contributions will be divided into sections containing that specific persons work.
## ---------- Import Libraries ---------------------
```{r}
library(tidyverse)
library(readr)
library(lubridate)
```

## ---------- Production Data Refinement ---------------------
```{r production_refinement}
# Import data, convert weeks into date datatype
production_data <- read_csv('CSV/weekly_factory_production.csv',
                            col_types = cols(
                              week = col_date(format = "%Y-%m-%d")))

# Are there null vals
colSums((is.na(production_data)))

# Add month column
production_data$MONTH_PRODUCED <- format(production_data$week, "%Y-%m")

# Calculate total production
production_data$total_units <- production_data$`Classic Cars` + production_data$Motorcycles + production_data$Planes + production_data$Ships + production_data$`Trucks and Buses`+production_data$`Vintage Cars`

# Calculate item production costs
production_data$`Weekly CPU` <- production_data$weekly_cost / production_data$total_units

# Aggregate the monthly data
monthly_data <- production_data %>%
  group_by(MONTH_PRODUCED) %>%
  summarise(
    MONTHLY_PRODUCTION = sum(total_units),
    MONTHLY_COST = sum(weekly_cost),
    MONTHLY_AVG_CPU = MONTHLY_COST / MONTHLY_PRODUCTION
  )
# Crate a long format of production data
long_production_data <- production_data %>% 
  group_by(MONTH_PRODUCED) %>%
  pivot_longer(
    cols = c(`Classic Cars`,Motorcycles, Planes, Ships, `Trucks and Buses`, `Vintage Cars`),
    names_to = "PRODUCTLINE",
    values_to = "UNITS_PRODUCED"
  )
# Calculate product costs
long_production_data$PRODUCT_COSTS <- long_production_data$`Weekly CPU`* long_production_data$UNITS_PRODUCED

# condense the data by month and product line
Condensed_production <- long_production_data %>%
  group_by(MONTH_PRODUCED, PRODUCTLINE) %>%
  summarise(
    TOTAL_UNITS_PRODUCED = sum(UNITS_PRODUCED),
    AVG_prod_cost = mean(PRODUCT_COSTS)
  )
# Add Cost per unit from monthly data set, manual calculations of compounded CPU lead to compounded floating point errors
Condensed_production <- Condensed_production %>%
  left_join(
    monthly_data %>% 
      select(MONTH_PRODUCED, MONTHLY_AVG_CPU),
    by = c("MONTH_PRODUCED" ="MONTH_PRODUCED")
  )
# Calculate production costs
Condensed_production$AVG_PRODUCTION_COST <- Condensed_production$MONTHLY_AVG_CPU * Condensed_production$TOTAL_UNITS_PRODUCED
  
summary(production_data)
glimpse(production_data)

```
## ---------- Sales Data Refinement ---------------------
```{r}
# Import sales data, Convert N/A values back to North America, and format date
sales_data <- read_csv("CSV/sales_data_sample_cleaned.csv")
sales_data$TERRITORY[is.na(sales_data$TERRITORY)] <- "NA"
sales_data$ORDERDATE <- mdy(sales_data$ORDERDATE)

# Create a Month_Ordered Column containing the month and year combined
sales_data$MONTH_ORDERED <- format(sales_data$ORDERDATE, "%Y-%m")

#Calculate average price by product type- used for product monthly data
sales_monthly_av_price <- sales_data %>%
  group_by(MONTH_ORDERED, PRODUCTLINE) %>%
  summarise(
      AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )

#Calculate average price by month - used for monthly data
Monthly_Sales_Data <- sales_data %>%
  group_by(MONTH_ORDERED) %>%
  summarise(
      AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )

```
## ---------- Combine Sales and Production ---------------------
```{r}
# Implement a Full join of the condensed production data and sales data
Production_and_Sales <- Condensed_production %>%
  dplyr::full_join(sales_monthly_av_price,
            by = c("MONTH_PRODUCED" = "MONTH_ORDERED",
                   "PRODUCTLINE" = "PRODUCTLINE"))
# Add Average Revenue per product
Production_and_Sales$AVG_REVENUE <- Production_and_Sales$TOTAL_UNITS_PRODUCED * Production_and_Sales$AVG_PRICE_SOLD
Production_and_Sales$AVG_PROFIT <- Production_and_Sales$AVG_REVENUE -
  Production_and_Sales$AVG_PRODUCTION_COST

# Implement a Full Join of monthly sales and monthly production
Monthly_Sales_and_Production <- Monthly_Sales_Data %>%
  dplyr::full_join(
    monthly_data,
    by = c('MONTH_ORDERED' = 'MONTH_PRODUCED')
  )
# Calculate Revenue
Monthly_Sales_and_Production$AVG_REVENUE <- Monthly_Sales_and_Production$MONTHLY_PRODUCTION * Monthly_Sales_and_Production$AVG_PRICE_SOLD

# Calculate Profit
Monthly_Sales_and_Production$AVG_PROFIT <- Monthly_Sales_and_Production$AVG_REVENUE - Monthly_Sales_and_Production$MONTHLY_AVG_CPU



pro <- Production_and_Sales %>%
  left_join(
    Monthly_Sales_and_Production %>% select(MONTH_ORDERED, MONTHLY_AVG_CPU),
    by = c("MONTH_PRODUCED" ="MONTH_ORDERED")
  )
pro$AVG_PRODUCTION_COST <- pro$Monthly_avg_CPU * pro$TOTAL_UNITS_PRODUCED

```




## ---------- Bea's Production ---------------------


## ---------- Eltons's Production ---------------------


## ---------- Kainen's Production ---------------------