---
title: "Project_3_Production"
author: "Bea Sauve, Elton nichols, Kainen Osborne"
date: "2026-03-03"
output: html_document
---

# Project 3 Production Notebook

This notebook will entail all team members work including data refinement, forecasting productions, and statistic generations. Team member contributions will be divided into sections containing that specific persons work. \## ---------- Import Libraries ---------------------

```{r}
library(tidyverse)
library(readr)       # Both readr and lubridate should be included already when installing tidyverse
library(lubridate)
library(forecast)
library(tseries)
```

## ---------- Production Data Refinement ---------------------

```{r production_refinement_Bea}
# Import data, convert weeks into date datatype
production_data <- read_csv('CSV/weekly_factory_production.csv',
                            col_types = cols(
                              week = col_date(format = "%Y-%m-%d")))
# Are there null vals
colSums((is.na(production_data)))

# Add month column
production_data$MONTH_PRODUCED <- format(production_data$week, "%Y-%m")

# Rename all columns to exclude spaces
production_data <- production_data %>%
  rename(
     "Classic_Cars" = `Classic Cars`,
     "Trucks_and_Buses" = `Trucks and Buses`,
     "Vintage_Cars" = `Vintage Cars` 
  )


# The outlier column indicates sections of weekly cost data that deviates from the norm. Create a flag column with 1 indicating that the row has outlieing data
production_data$outlier_flag <- if_else(production_data$outlier > 0, 1, 0)

# Calculate total production
production_data$TOTAL_UNITS <- production_data$Classic_Cars + production_data$Motorcycles + production_data$Planes + production_data$Ships + production_data$Trucks_and_Buses+production_data$Vintage_Cars

# Calculate item production costs
production_data$`AGGREGATE_CPU` <- production_data$weekly_cost / production_data$TOTAL_UNITS
```

##-----------Sales Data implementation------------

```{r}
# Import sales data, Convert N/A values back to North America, and format date
sales_data <- read_csv("CSV/sales_data_sample_cleaned.csv")
sales_data$TERRITORY[is.na(sales_data$TERRITORY)] <- "NA"
sales_data$ORDERDATE <- mdy(sales_data$ORDERDATE)

# Create a Month_Ordered Column containing the month and year combined
sales_data$MONTH_ORDERED <- format(sales_data$ORDERDATE, "%Y-%m")

sales_data <- sales_data %>%
  mutate(PRODUCTLINE = gsub(" ", "_", PRODUCTLINE))

```

## ---------------Regression Model for Productline Prices----------------

We would like to determine the actual cost per unit not based on aggregated data. A multiple linear regression model will assist in the isolation of each product cost under consideration that labor, cost of electricity, etc are consistent.

```{r product_regression_Bea}
regression_products <- lm(data = production_data, weekly_cost ~ Classic_Cars + Motorcycles +
                            Planes + Ships + Trucks_and_Buses + Vintage_Cars)
summary(regression_products)
```

Testing the models by factory

```{r production_by_factory_Bea}
F0_Production <- production_data %>% filter(factory_id == "0")
F1_Production <- production_data %>% filter(factory_id == "1")
F2_Production <- production_data %>% filter(factory_id == "2")

```

```{r factory_regression_bea}
f0_regression <- lm(data = F0_Production, weekly_cost ~ Classic_Cars + Ships)
summary(f0_regression)

f1_regression <- lm(data = F1_Production,weekly_cost ~  Motorcycles +Trucks_and_Buses)
summary(f1_regression)

f2_regression <- lm(data = F2_Production,weekly_cost ~ Planes+Vintage_Cars)
summary(f2_regression)

#include the factory for the regression
Master_Regression <- lm(weekly_cost ~ Classic_Cars + Motorcycles + Planes + Ships + 
                  Trucks_and_Buses + Vintage_Cars + factor(factory_id), 
                  data = production_data,
                  weights = if_else(production_data$outlier == 0, 1, 0))
summary(Master_Regression)
par(mfrow=c(2,2))
plot(Master_Regression)
par(mfrow=c(1,1))
```

## -----------Convert dataset into long format---------------------

```{r production_long_format_Bea}
# Crate a long format of production data
weekly_productions_by_productline <- production_data %>% 
  group_by(MONTH_PRODUCED) %>%
  pivot_longer(
    cols = c(Classic_Cars,Motorcycles, Planes, Ships, Trucks_and_Buses, Vintage_Cars),
    names_to = "PRODUCTLINE",
    values_to = "UNITS_PRODUCED"
  )
# Calculate product costs
weekly_productions_by_productline$AGG_PRODUCT_COSTS <- weekly_productions_by_productline$`AGGREGATE_CPU`* weekly_productions_by_productline$UNITS_PRODUCED
 
# Assign predicted coefficients of regression model to long format model
reg_coeffs <- coef(Master_Regression)
#Clean up the back ticks generated in production line names
names(reg_coeffs) <- gsub("`","", names(reg_coeffs))
# Define the base production intercept
base_intercept <- reg_coeffs["(Intercept)"]

# Implement regression calculated product cost per unit
weekly_productions_by_productline <- weekly_productions_by_productline %>%
  mutate(
    REGRESSION_CPU = ifelse(TOTAL_UNITS > 0,
                            reg_coeffs[PRODUCTLINE],
                            0)
    )

# add intercepts based on factory id_ factory 0 is base ID
weekly_productions_by_productline <- weekly_productions_by_productline %>%
  mutate(
    REG_FACTORY_BASE_COST = base_intercept + case_when(
      factory_id == 1 ~ reg_coeffs["factor(factory_id)1"],
      factory_id == 2 ~ reg_coeffs["factor(factory_id)2"],
      TRUE            ~ 0 # This handles Factory 0
    )
  )
# add regression product costs
weekly_productions_by_productline$REG_TTL_PROD_COST_NO_BASE <- weekly_productions_by_productline$REGRESSION_CPU * weekly_productions_by_productline$UNITS_PRODUCED

weekly_productions_by_productline
```

## -------------Summarize weekly production into Monthly-------------------------

Condense the data set into a monthly data set instead of a weekly dataset. Implement regression model information into the dataset

```{r condensed_regression_data_Bea}
# condense the data by month and product line
monthly_productions_factory_X_productline <- weekly_productions_by_productline %>%
  group_by(MONTH_PRODUCED, PRODUCTLINE, factory_id) %>%
  summarise(
    TOTAL_UNITS_PRODUCED = sum(UNITS_PRODUCED),
    .groups = "drop"
  ) %>% filter(TOTAL_UNITS_PRODUCED != 0)
# Implement regression calculated product cost per unit
monthly_productions_factory_X_productline <- monthly_productions_factory_X_productline %>%
  mutate(
    REGRESSION_CPU = ifelse(TOTAL_UNITS_PRODUCED > 0,
                            reg_coeffs[PRODUCTLINE],
                            0)
    )

# add intercepts based on factory id_ factory 0 is base ID
monthly_productions_factory_X_productline <-monthly_productions_factory_X_productline %>%
  mutate(
    REG_FACTORY_BASE_COST = base_intercept + case_when(
      factory_id == 1 ~ reg_coeffs["factor(factory_id)1"],
      factory_id == 2 ~ reg_coeffs["factor(factory_id)2"],
      TRUE            ~ 0 # This handles Factory 0
    )
  )
# add regression product costs without base
monthly_productions_factory_X_productline$REG_PROD_COST_NO_BASE <-monthly_productions_factory_X_productline$REGRESSION_CPU * monthly_productions_factory_X_productline$TOTAL_UNITS_PRODUCED
# add regression product costs with base
monthly_productions_factory_X_productline$AVG_PRODUCTION_COST_W_BASE <- monthly_productions_factory_X_productline$REGRESSION_CPU *
  monthly_productions_factory_X_productline$TOTAL_UNITS_PRODUCED + 
  monthly_productions_factory_X_productline$REG_FACTORY_BASE_COST

monthly_productions_factory_X_productline

```

#----------AGGREGATE MONTLHY DATA----------- Note: Regression monthly calculations could be refined more

```{r Monthly_refinement_Bea}
#Calculate average price by product type- used for product monthly data
monthly_sales_by_productline <- sales_data %>%
  group_by(MONTH_ORDERED, PRODUCTLINE) %>%
  summarise(
    MIN_PRICE_SOLD = min(PRICEEACH, na.rm = TRUE),
      AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    MAX_PRICE_SOLD = max(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )
monthly_sales_by_productline

Final_SalesXProduction_Productline <- monthly_productions_factory_X_productline %>%
  full_join(monthly_sales_by_productline,
            by = c("MONTH_PRODUCED" = "MONTH_ORDERED","PRODUCTLINE"))
# Calculate profits per product
Final_SalesXProduction_Productline$MIN_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$MIN_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
Final_SalesXProduction_Productline$AVG_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$AVG_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
Final_SalesXProduction_Productline$Max_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$MAX_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
            
# Add Average Revenue per product
Final_SalesXProduction_Productline$MIN_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$MIN_PRICE_SOLD
Final_SalesXProduction_Productline$AVG_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$AVG_PRICE_SOLD
Final_SalesXProduction_Productline$MAX_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$MAX_PRICE_SOLD
# Add average profit 
Final_SalesXProduction_Productline$MIN_TOTAL_PROFIT <- Final_SalesXProduction_Productline$MIN_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$AVG_TOTAL_PROFIT <- Final_SalesXProduction_Productline$AVG_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$MAX_TOTAL_PROFIT <- Final_SalesXProduction_Productline$MAX_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$ACTUAL_TOTAL_PROFIT <- Final_SalesXProduction_Productline$TOTAL_SALES - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE

Final_SalesXProduction_Productline
```

Aggregate monthly sales no product line

```{r}
#predicted weekly costs
production_data$PREDICTED_WEEKLY_COST <- predict(Master_Regression, newdata = production_data)

Summary_Monthly_Data <- production_data %>%
  group_by(MONTH_PRODUCED) %>%
  summarise(
    MONTHLY_PRODUCTION  = sum(TOTAL_UNITS),
    ACTUAL_MONTHLY_COST = sum(weekly_cost),
    PRED_MONTHLY_COST   = sum(PREDICTED_WEEKLY_COST),
    #Actual Cost Per Unit
    ACTUAL_AVG_CPU      = ACTUAL_MONTHLY_COST / MONTHLY_PRODUCTION,
    # Predicted Cost Per Unit
    REGRESSION_AVG_CPU  = PRED_MONTHLY_COST / MONTHLY_PRODUCTION,
    CPU_VARIANCE        = ACTUAL_AVG_CPU - REGRESSION_AVG_CPU
  )
#Calculate average price by month - used for monthly data
Monthly_Sales_Data <- sales_data %>%
  group_by(MONTH_ORDERED) %>%
  summarise(
    MIN_PRICE_SOLD = min(PRICEEACH, na.rm = TRUE),
    AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    MAX_PRICE_SOLD = max(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )
# FUll join of sales and production
Final_SalesXProduction_Month <- Summary_Monthly_Data %>%
  full_join(Monthly_Sales_Data,
            by = c("MONTH_PRODUCED" = "MONTH_ORDERED"))
# Ensure Dates are date data types
Final_SalesXProduction_Productline$Month <- ym(Final_SalesXProduction_Productline$MONTH_PRODUCED)

# Calculate profits per product
Final_SalesXProduction_Month$MIN_PRODUCT_PROFIT <- Final_SalesXProduction_Month$MIN_PRICE_SOLD - Final_SalesXProduction_Month$REGRESSION_AVG_CPU
Final_SalesXProduction_Month$AVG_PRODUCT_PROFIT <- Final_SalesXProduction_Month$AVG_PRICE_SOLD - Final_SalesXProduction_Month$REGRESSION_AVG_CPU
Final_SalesXProduction_Month$Max_PRODUCT_PROFIT <- Final_SalesXProduction_Month$MAX_PRICE_SOLD - Final_SalesXProduction_Month$REGRESSION_AVG_CPU

#add Revenue
Final_SalesXProduction_Month$MIN_TOTAL_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$MIN_PRICE_SOLD
Final_SalesXProduction_Month$AVG_TOTAL_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$AVG_PRICE_SOLD
Final_SalesXProduction_Month$MAX_TOTAL_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$MAX_PRICE_SOLD

#add profit
Final_SalesXProduction_Month$MIN_TOTAL_PROFIT <- Final_SalesXProduction_Month$MIN_TOTAL_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$AVG_TOTAL_PROFIT <- Final_SalesXProduction_Month$AVG_TOTAL_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$MAX_TOTAL_PROFIT <- Final_SalesXProduction_Month$MAX_TOTAL_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$ACTUAL_TOTAL_PROFIT <- Final_SalesXProduction_Month$TOTAL_SALES - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
```

## ----------Export the Datasets for Visualizing-----------

```{r}
# write_csv(Final_SalesXProduction_Productline, file.choose(new = TRUE))
# write_csv(Final_SalesXProduction_Month, file.choose(new = TRUE))
```

## ---------- Bea's Production ---------------------

Forcast expected productline sales from june 2005 - December 2005

```{r}

# Group the data by date and product for time series
monthly_data <- sales_data %>%
  group_by(MONTH_ORDERED, PRODUCTLINE) %>%
  summarise(TOTAL_UNITS = sum(QUANTITYORDERED),
            .groups = "drop") %>%
   arrange(PRODUCTLINE, MONTH_ORDERED)

monthly_data$MONTH_ORDERED <- ym(monthly_data$MONTH_ORDERED)
  
#--------------Determine Which method to use per production line------------
production_model <- function(product_line) {
  # Create a productline time table
  submodel <- monthly_data %>%
    filter(PRODUCTLINE == product_line) %>%
    arrange(MONTH_ORDERED)
  
  # Build Time Series object
  ts_submodel <-ts(submodel$TOTAL_UNITS,
                   start = c(year(min(submodel$MONTH_ORDERED)),
                             month(min(submodel$MONTH_ORDERED))),
                   frequency = 12)
  last_date <- max(submodel$MONTH_ORDERED)
  target_end <- as.Date("2005-12-31")
  h <- interval(last_date, target_end) %/% months(1)
  
  mean_model <- meanf(ts_submodel, h = h)
  snaive_model <- snaive(ts_submodel, h = h)
  rwf_model <- rwf(ts_submodel, h = h, drift = TRUE)
  
  ets_ann <- ets(ts_submodel, model = "ANN")
  forecast_ets_ann <- forecast(ets_ann, h = h)
  
  ets_aaa <- ets(ts_submodel, model = "AAA")
  forecast_ets_aaa <- forecast(ets_aaa, h = h)
  
  ets_auto <- ets(ts_submodel)
  forecast_ets_auto <- forecast(ets_auto, h = h)
  
  arima_auto <- auto.arima(ts_submodel)
  forecast_arima <- forecast(arima_auto, h = h)
  
  lm_model <- tslm(ts_submodel ~ trend + season)
  forecast_lm <- forecast(lm_model, h = h)
  
  # STL needs 2 full years - safely skip if not enough data
  stl_forecast <- tryCatch(
    stlf(ts_submodel, h = h),
    error = function(e) NULL
  )
  
  accuracy_table <- list(
    Mean = accuracy(mean_model)[1,],
    Seasonal_Naive = accuracy(snaive_model)[1,],
    Random_Walk_Drift = accuracy(rwf_model)[1,],
    ETS_ANN = accuracy(forecast_ets_ann)[1,],
    ETS_AAA = accuracy(forecast_ets_aaa)[1,],
    ETS_Auto = accuracy(forecast_ets_auto)[1,],
    ARIMA = accuracy(forecast_arima)[1,],
    Linear_Regression = accuracy(forecast_lm)[1,]
  )
  if (!is.null(stl_forecast)) {
    accuracy_table$STL <- accuracy(stl_forecast)[1,]
  }
  
  accuracy_table <- round(do.call(rbind, accuracy_table), 2)
  cat("\n===", product_line, "===\n")
  print(accuracy_table)
  
}

#Run all forecasts
product_lines <- unique(monthly_data$PRODUCTLINE)
results <- map(product_lines, production_model)
results <- results[!sapply(results, is.null)]
```

#### Orders per product line Time series

```{r}
forecast_production <- function(product_line) {
  submodel <- monthly_data %>%
    filter(PRODUCTLINE == product_line) %>%
    arrange(MONTH_ORDERED) %>%
    # Fill in missing order points with 0's
    tidyr::complete(
      MONTH_ORDERED = seq.Date(min(MONTH_ORDERED), max(MONTH_ORDERED), by = "month"),
      fill = list(TOTAL_UNITS = 0)
    )
  
  ts_submodel <- ts(submodel$TOTAL_UNITS,
                 start = c(year(min(submodel$MONTH_ORDERED)),
                           month(min(submodel$MONTH_ORDERED))),
                 frequency = 12)
  # Calculate Time series ranges
  last_date <- max(submodel$MONTH_ORDERED)
  target_end <- as.Date("2005-12-31")
  h <- interval(last_date, target_end) %/% months(1)
  
  # Use different series depending num months in dataset
  if (length(ts_submodel) < 24){
    ts_forecast <- forecast(tslm(ts_submodel ~ trend + season), h = h) 
  } else{
    stl_model <- stl(ts_submodel, s.window = "periodic")
    ts_forecast <- forecast(stl_model, h=h)
  }
  # Display forecasting data
  cat("\n===", product_line, "===\n")
  print(ts_forecast)
  
  #plot the forecast
  autoplot(ts_forecast) +
    ggtitle(paste(product_line, "Orders Forecast"))
  
}

#Run all forecasts
product_lines <- unique(monthly_data$PRODUCTLINE)
#Map the products to the function
results <- map(product_lines, forecast_production)
# Omit null values
results <- results[!sapply(results, is.null)]
print(results)
```

## ---------- Elton's Production ---------------------

### EDA

#### Production Trends Over Time

```{r Load and Sort Data}
# Loading the data
df_total <- read_csv("monthly_salesXproduction.csv", show_col_types = FALSE)
df_line <- read_csv("monthly_salesXproduction_by_productline.csv", show_col_types = FALSE)

# Sort chronologically by month
df_total <- df_total %>% arrange(MONTH_PRODUCED)
df_line <- df_line %>% arrange(MONTH_PRODUCED)
```

```{r Total Monthly Production Over Time}
plot_total <- ggplot(df_total, aes(x = MONTH_PRODUCED, y = MONTHLY_PRODUCTION)) +
  geom_line(color = "blue", group = 1) +
  geom_point(color = "blue") +
  labs(title = "Total Monthly Production Over Time",
       x = "Month Produced",
       y = "Total Units Produced") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # Making x-axis labels more readable
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), 
                      by = max(1, floor(length(x)/15)))])
print(plot_total)
```

```{r Monthly Production by Product Line Over Time, echo = FALSE}
df_line_filtered <- df_line %>% 
  drop_na(factory_id) %>% 
  mutate(factory_label = paste("Factory", factory_id),
         MONTH_DATE = ym(MONTH_PRODUCED))

plot_productline <- ggplot(df_line_filtered, aes(x = MONTH_DATE, y = TOTAL_UNITS_PRODUCED,
                                        color = PRODUCTLINE, group = PRODUCTLINE)) +
  geom_line() +
  geom_point() +
  labs(title = "Monthly Production by Product Line Over Time",
       x = "Month Produced",
       y = "Units Produced",
       color = "Product Line") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_breaks = "2 months", date_labels = "%b `%y")

print(plot_productline)
```

```{r Monthly Production by Product Line Over Time (Faceted by Factory), echo = FALSE}
# Making cleaner labels
df_line_filtered <- df_line %>% 
  # Removing the factory with no data
  drop_na(factory_id) %>% 
  mutate(factory_label = paste("Factory", factory_id),
         MONTH_DATE = ym(MONTH_PRODUCED))

plot_productline_faceted <- ggplot(df_line_filtered, aes(x = MONTH_DATE, y = TOTAL_UNITS_PRODUCED,
                                        color = PRODUCTLINE, group = PRODUCTLINE)) +
  geom_line() +
  geom_point() +
  # Splitting the plot by factory
  facet_wrap(~ factory_label) +
  labs(title = "Monthly Production by Product Line Over Time (By Factory)",
       x = "Month Produced",
       y = "Units Produced",
       color = "Product Line") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        strip.text = element_text(face = "bold")) +
  # Easier to read label formatting, every four months for space savings
  scale_x_date(date_breaks = "4 months", date_labels = "%b `%y")

print(plot_productline_faceted)
```

#### Seasonality / Trend Patterns

```{r Time Series Decomposition for Seasonal Patterns and Trends}
# Making a column of Date objects
df_total = df_total %>% 
  mutate(MONTH_DATE = ym(MONTH_PRODUCED))

# Converting to a Time Series object
start_year <- year(min(df_total$MONTH_DATE))
start_month <- month(min(df_total$MONTH_DATE))

# Frequency of 12 for 12 months
production_ts <- ts(df_total$MONTHLY_PRODUCTION,
                    start = c(start_year, start_month),
                    frequency = 12)

# STL to split into Seasonality, Trend, and Remainder
decomposition <- stl(production_ts, s.window = "periodic")
plot_decomp <- autoplot(decomposition) +
  theme_minimal() +
  labs(title = "Time Series Decomposition of Total Monthly Production")

print(plot_decomp)
```

```{r Time Series Decomposition by Product Line}
# Ignoring product lines with no production, i.e. Trains
valid_product_lines <- df_line %>% 
  group_by(PRODUCTLINE) %>% 
  summarize(total = sum(TOTAL_UNITS_PRODUCED, na.rm = TRUE)) %>% 
  filter(total > 0) %>% 
  pull(PRODUCTLINE)

# Loop through each valid product line to generate plots
for (prod_line in valid_product_lines) {
  # Filter and prepare data for the specific product line
  df_sub <- df_line %>% 
    filter(PRODUCTLINE == prod_line) %>% 
    # If a month is missing, fill it with 0
    complete(MONTH_PRODUCED = unique(df_line$MONTH_PRODUCED),
             fill = list(TOTAL_UNITS_PRODUCED = 0)) %>% 
    arrange(MONTH_PRODUCED)
  
  # Create a Time Series object
  prod_ts <- ts(df_sub$TOTAL_UNITS_PRODUCED,
                start = c(start_year, start_month),
                frequency = 12)
  # Plot Decomposition
  decomp <- stl(prod_ts, s.window = "periodic")
  plot_decomp <- autoplot(decomp) +
    theme_minimal() +
    labs(title = paste("Time Series Decomposition of Monthly Production:", prod_line))
  
  print(plot_decomp)
  
  # Plot Seasonal Plot
  plot_season <- ggseasonplot(prod_ts, year.labels = TRUE, continuous = TRUE) +
  theme_minimal() +
  labs(title = paste("Seasonal Plot:", prod_line),
       y = "Units Produced",
       x = "Month")
  
  print(plot_season)
}
```

#### Descriptive Statistics

-   Factory

```{r}

```

-   Product Line

-   Month

```{r Monthly Production Statistics}
# Calculating average monthly production
cat("Average Monthly Production:\t", round(mean(df_total$MONTHLY_PRODUCTION), 2), "units\n")

# Find peak production month
peak_month <- df_total %>% filter(MONTHLY_PRODUCTION == max(MONTHLY_PRODUCTION))
cat("Peak Production Month:\t\t", peak_month$MONTH_PRODUCED, "with", peak_month$MONTHLY_PRODUCTION, "units\n")

# Find lowest production month
lowest_month <- df_total %>% filter(MONTHLY_PRODUCTION == min(MONTHLY_PRODUCTION))
cat("Lowest Production Month:\t", lowest_month$MONTH_PRODUCED, "with", lowest_month$MONTHLY_PRODUCTION, "units\n")

cat("Total Production by Product Line:\n")
product_line_summary <- df_line %>% 
  group_by(PRODUCTLINE) %>% 
  summarize(TOTAL_PRODUCED = sum(TOTAL_UNITS_PRODUCED, na.rm = TRUE)) %>% 
  arrange(desc(TOTAL_PRODUCED))

print(product_line_summary)
```

#### Discussion

-   Are trends visible?

    -   Yes, overall there is a strong upward trend for the first year and four months, followed by a leveling out and then small decrease.

-   Are there seasonal effects?

    -   All of the product lines show perfectly repeating seasonal patterns.

-   Do product lines differ substantially?

## ---------- Kainen's Production ---------------------

```{r kainen_setup}
suppressPackageStartupMessages({
  library(forecast)
  library(tseries)
  library(scales)
  library(gridExtra)
})

# Add date column to joined product-line dataframe
Final_SalesXProduction_Productline$date <- as.Date(
  paste0(Final_SalesXProduction_Productline$MONTH_PRODUCED, "-01")
)

# Color palette used across all of Kainen's plots
# Codes are taken from online.
PALETTE <- c(
  "Classic_Cars"     = "#E63946",
  "Motorcycles"      = "#457B9D",
  "Planes"           = "#2A9D8F",
  "Ships"            = "#E9C46A",
  "Trucks_and_Buses" = "#6A0572",
  "Vintage_Cars"     = "#264653"
)

# Product lines with tracked production
product_lines <- c("Classic_Cars", "Motorcycles", "Planes",
                   "Ships", "Trucks_and_Buses", "Vintage_Cars")

forecast_horizon <- 7
```

### | Monthly Time Series Objects

This converts the cleaned product-line dataframe into one time series object per
product line.


```{r kainen_build_ts}
ts_list <- list()

for (pl in product_lines) {
  pl_data <- Final_SalesXProduction_Productline %>%
    filter(PRODUCTLINE == pl, !is.na(TOTAL_UNITS_PRODUCED)) %>%
    arrange(date)

  if (nrow(pl_data) == 0) {
    cat(sprintf("%-20s  SKIPPED -- no production data\n", pl))
    next
  }

  ts_list[[pl]] <- ts(
    pl_data$TOTAL_UNITS_PRODUCED,
    start     = c(year(min(pl_data$date)), month(min(pl_data$date))),
    frequency = 12
  )
}

# Update product_lines to only those with a valid ts object
product_lines <- names(ts_list)

# Print coverage for each series
for (pl in product_lines) {
  cat(sprintf("%-20s  start = %s  end = %s  n = %d obs\n",
              pl,
              paste(start(ts_list[[pl]]), collapse = "-"),
              paste(end(ts_list[[pl]]),   collapse = "-"),
              length(ts_list[[pl]])))
}
```



### | Overall Production of each Product Line.

```{r kainen_plot_series, fig.height=10, fig.width=12}
series_plots <- list()

for (pl in product_lines) {
  series_plots[[pl]] <- autoplot(ts_list[[pl]]) +
    ggtitle(pl) +
    xlab("Year") +
    ylab("Units Produced") +
    theme_minimal()
}

do.call(grid.arrange, c(series_plots, list(ncol = 2,
        top  = "Monthly Units Produced by Product Line (Jan 2003 - May 2005)")))
```


### | Stationarity and Seasonality Checks


```{r kainen_ndiffs}
# Checking for difference
for (pl in product_lines) {
  d <- ndiffs(ts_list[[pl]])
  cat(sprintf("%-20s  ndiffs = %d\n", pl, d))
}
```

### | Augmented Dickey-Fuller test

The ADF test checks whether the series has a unit root (i.e. is
non-stationary).

If p < 0.05 we have stationarity.
If p >= 0.05 there is evidence of non-stationarity.

```{r kainen_adf}
for (pl in product_lines) {
  result <- adf.test(ts_list[[pl]])
  cat(sprintf("%-20s  ADF p-value = %.4f  (%s)\n",
              pl,
              result$p.value,
              ifelse(result$p.value < 0.05, "stationary", "non-stationary")))
}
```

### | Seasonality detection via STL

```{r kainen_seasonality}
for (pl in product_lines) {
  ts_obj  <- ts_list[[pl]]
  stl_fit <- stl(ts_obj, s.window = "periodic")
  sr <- var(stl_fit$time.series[, "seasonal"]) / var(ts_obj)
  cat(sprintf("%-20s  seasonal ratio = %.3f  (%s)\n",
              pl,
              sr,
              ifelse(sr > 0.10, "seasonality detected", "weak or no seasonality")))
}
```

### | Comparing All Four Model Types

```{r kainen_fit_models}
fit_store    <- list()
model_labels <- list()
aicc_table   <- data.frame()

for (pl in product_lines) {
  ts_obj <- ts_list[[pl]]
  n      <- length(ts_obj)

  # -- ARIMA --
  fit_arima <- tryCatch(
    auto.arima(ts_obj, stepwise = FALSE, approximation = FALSE, ic = "aicc"),
    error = function(e) NULL
  )

  # -- ETS --
  fit_ets <- tryCatch(
    ets(ts_obj, ic = "aicc"),
    error = function(e) NULL
  )

  # -- STL (ETS on seasonally adjusted component) --
  fit_stl <- tryCatch(
    stlf(ts_obj, h = forecast_horizon),
    error = function(e) NULL
  )

  # -- Linear regression with trend + season --
  fit_lm <- tryCatch(
    tslm(ts_obj ~ trend + season),
    error = function(e) NULL
  )

  # Collect AICc values (STL and LM use AIC as proxy)
  aicc_arima <- if (!is.null(fit_arima)) fit_arima$aicc   else NA
  aicc_ets   <- if (!is.null(fit_ets))   fit_ets$aicc     else NA
  aicc_stl   <- if (!is.null(fit_stl))   AIC(fit_stl$model) else NA
  aicc_lm    <- if (!is.null(fit_lm))    AIC(fit_lm)      else NA

  aicc_table <- rbind(aicc_table, data.frame(
    product_line = pl,
    ARIMA        = round(aicc_arima, 1),
    ETS          = round(aicc_ets,   1),
    STL          = round(aicc_stl,   1),
    LM           = round(aicc_lm,    1)
  ))

  # Select best model by lowest AICc
  candidates <- c(ARIMA = aicc_arima, ETS = aicc_ets, STL = aicc_stl, LM = aicc_lm)
  best_name  <- names(which.min(candidates))

  best_fit <- switch(best_name,
    ARIMA = fit_arima,
    ETS   = fit_ets,
    STL   = fit_stl$model,
    LM    = fit_lm
  )

  fit_store[[pl]]    <- list(arima = fit_arima, ets = fit_ets,
                             stl   = fit_stl,   lm  = fit_lm,
                             best  = best_fit,  best_name = best_name)
  model_labels[[pl]] <- best_name

  cat(sprintf("\n%s -- selected: %s  (AICc: ARIMA=%.1f  ETS=%.1f  STL=%.1f  LM=%.1f)\n",
              pl, best_name,
              aicc_arima, aicc_ets, aicc_stl, aicc_lm))
}
```

### | AICc Summary Table

```{r kainen_aicc_table}
print(aicc_table, row.names = FALSE)
```

### | Accuracy Comparison

- **RMSE** -- root mean squared error. Penalizes large errors more heavily.
- **MAE** -- mean absolute error. Average absolute deviation from true values.
- **MAPE** -- mean absolute percentage error. Error as a percentage of true value,
  easier to interpret across series with different scales.
- **MASE** -- compares the model against the seasonal naive baseline. Values below
  1.0 mean the model beats simply repeating last year's value.
- **ACF1** -- autocorrelation of residuals at lag 1. Values near zero mean the
  model has captured the serial structure and left behind noise.

```{r kainen_accuracy, warning=FALSE}
for (pl in product_lines) {
  ts_obj <- ts_list[[pl]]
  fits   <- fit_store[[pl]]

  fc_arima <- if (!is.null(fits$arima)) forecast(fits$arima, h = forecast_horizon) else NULL
  fc_ets   <- if (!is.null(fits$ets))   forecast(fits$ets,   h = forecast_horizon) else NULL
  fc_stl   <- fits$stl
  fc_lm    <- if (!is.null(fits$lm))    forecast(fits$lm,    h = forecast_horizon) else NULL

  rows <- list()
  if (!is.null(fc_arima)) rows[["ARIMA"]] <- accuracy(fc_arima)[1, ]
  if (!is.null(fc_ets))   rows[["ETS"]]   <- accuracy(fc_ets)[1, ]
  if (!is.null(fc_stl))   rows[["STL"]]   <- accuracy(fc_stl)[1, ]
  if (!is.null(fc_lm))    rows[["LM"]]    <- accuracy(fc_lm)[1, ]

  if (length(rows) > 0) {
    tbl <- round(do.call(rbind, rows), 2)
    cat(sprintf("\n--- %s (selected: %s) ---\n", pl, model_labels[[pl]]))
    print(tbl)
  }
}
```

### | Model Diagnostics

```{r kainen_diagnostics, fig.height=5, fig.width=10}
for (pl in product_lines) {
  fits      <- fit_store[[pl]]
  best_name <- model_labels[[pl]]
  best_fit  <- fits$best

  cat(sprintf("\n=== %s -- diagnostics for selected model: %s ===\n", pl, best_name))
  checkresiduals(best_fit)
}
```



### | Forecasts for June to December 2005

This step generates a 7-month forecast for each product line using the selected
model. Each plot shows:

- **Solid black line** : the historical production series (Jan 2003 - May 2005)
- **Solid Blue line** : the point forecast for Jun - Dec 2005
- **Dark shaded band** : 80% prediction interval
- **Light shaded band** : 95% prediction interval

```{r kainen_forecasts, fig.height=5, fig.width=10}
fc_store <- list()

for (pl in product_lines) {
  fits      <- fit_store[[pl]]
  best_name <- model_labels[[pl]]
  fc <- fits$stl
  fc_store[[pl]] <- fc
  print(
    autoplot(fc) +
      coord_cartesian(xlim = c(2005, 2006), ylim = c(0, 250)) +
      scale_x_continuous(
        breaks = 2005 + (0:11) / 12,
        labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                   "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
      ) +
      ggtitle(paste(pl, "-- forecast Jun-Dec 2005 (", best_name, ")")) +
      xlab("Month (2005)") +
      ylab("Units Produced") +
      theme_minimal()
  )
}
```


### | Average Profit Per Unit


```{r kainen_ppu}
# Computed in line
profit_per_unit <- Final_SalesXProduction_Productline %>%
  filter(!is.na(TOTAL_SALES),
         !is.na(AVG_PRODUCTION_COST_W_BASE),
         !is.na(TOTAL_UNITS_PRODUCED),
         TOTAL_UNITS_PRODUCED > 0) %>%
  mutate(
    actual_profit = TOTAL_SALES - AVG_PRODUCTION_COST_W_BASE,
    ppu           = actual_profit / TOTAL_UNITS_PRODUCED
  ) %>%
  group_by(PRODUCTLINE) %>%
  summarise(avg_profit_per_unit = mean(ppu, na.rm = TRUE),
            n_months = n(),
            .groups = "drop") %>%
  rename(product_line = PRODUCTLINE) %>%
  arrange(desc(avg_profit_per_unit))

print(profit_per_unit, digits = 2)
```

### | Profit Projections for June to December 2005

```{r kainen_profit_calc}
# Build a dataframe of point forecasts for Jun-Dec 2005
forecast_months <- seq(as.Date("2005-06-01"), as.Date("2005-12-01"), by = "month")

fc_rows <- list()
for (pl in product_lines) {
  fc     <- fc_store[[pl]]
  fc_rows[[pl]] <- data.frame(
    product_line = pl,
    month        = forecast_months,
    forecast     = pmax(0, as.numeric(fc$mean)),
    lower80      = pmax(0, as.numeric(fc$lower[, 1])),
    upper80      = pmax(0, as.numeric(fc$upper[, 1]))
  )
}

fc_df <- do.call(rbind, fc_rows)

# Join profit per unit and calculate projected profit
fc_profit <- merge(fc_df,
                   profit_per_unit[, c("product_line", "avg_profit_per_unit")],
                   by = "product_line")

fc_profit$projected_profit <- fc_profit$forecast   * fc_profit$avg_profit_per_unit
fc_profit$profit_lo80      <- fc_profit$lower80    * fc_profit$avg_profit_per_unit
fc_profit$profit_hi80      <- fc_profit$upper80    * fc_profit$avg_profit_per_unit
fc_profit$month_label      <- format(fc_profit$month, "%b %Y")

# Summary by product line
profit_by_pl <- aggregate(
  cbind(total_units = forecast,
        total_profit = projected_profit,
        lower_80 = profit_lo80,
        upper_80 = profit_hi80) ~ product_line,
  data = fc_profit, FUN = sum
)
profit_by_pl <- profit_by_pl[order(-profit_by_pl$total_profit), ]

# Factory map
factory_map <- unique(Final_SalesXProduction_Productline[
  !is.na(Final_SalesXProduction_Productline$factory_id),
  c("PRODUCTLINE", "factory_id")
])
names(factory_map)[1] <- "product_line"
factory_map <- factory_map[!duplicated(factory_map$product_line), ]

profit_by_factory <- merge(fc_profit, factory_map, by = "product_line", all.x = TRUE)
profit_by_factory <- aggregate(
  cbind(total_profit = projected_profit,
        lower_80 = profit_lo80,
        upper_80 = profit_hi80) ~ factory_id,
  data = profit_by_factory[!is.na(profit_by_factory$factory_id), ],
  FUN = sum
)

overall_profit <- sum(profit_by_pl$total_profit, na.rm = TRUE)

cat("\n=== Projected Profit by Product Line (Jun-Dec 2005) ===\n")
profit_by_pl$total_profit_fmt <- dollar(round(profit_by_pl$total_profit))
profit_by_pl$lower_80_fmt     <- dollar(round(profit_by_pl$lower_80))
profit_by_pl$upper_80_fmt     <- dollar(round(profit_by_pl$upper_80))
print(profit_by_pl[, c("product_line", "total_units",
                        "total_profit_fmt", "lower_80_fmt", "upper_80_fmt")],
      row.names = FALSE)

cat(sprintf("\nCompany Total (excl. Trains): %s\n", dollar(round(overall_profit))))

cat("\n=== Projected Profit by Factory (Jun-Dec 2005) ===\n")
print(profit_by_factory, row.names = FALSE)
```

### | Profit Projection Plots

#### # Plot 1: Monthly Stacked Profit

This plot shows projected profit for each month from June to December 2005,
broken down by product line.

```{r kainen_plot_monthly_profit, fig.height=5, fig.width=10}
month_order <- format(forecast_months, "%b %Y")
fc_profit$month_label <- factor(fc_profit$month_label, levels = month_order)

p_monthly <- ggplot(fc_profit,
                    aes(x = month_label,
                        y = projected_profit / 1e6,
                        fill = product_line)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = PALETTE) +
  scale_y_continuous(labels = dollar_format(suffix = "M", prefix = "$")) +
  labs(title   = "Projected Monthly Profit by Product Line (Jun-Dec 2005)",
       x       = NULL,
       y       = "Projected Profit (USD Millions)",
       fill    = "Product Line",
       caption = "Point forecast x average historical profit per unit") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

p_monthly
```

#### # Plot 2: Total Profit by Product Line

This bar chart ranks each product line by its total projected profit across the
full June to December 2005 period.

```{r kainen_plot_total_profit, fig.height=5, fig.width=9}
profit_by_pl$product_line <- reorder(profit_by_pl$product_line,
                                     profit_by_pl$total_profit)

p_total <- ggplot(profit_by_pl,
                  aes(x = total_profit / 1e6,
                      y = product_line,
                      fill = as.character(product_line))) +
  geom_col(width = 0.6, show.legend = FALSE) +
  scale_fill_manual(values = PALETTE) +
  scale_x_continuous(labels = dollar_format(suffix = "M", prefix = "$")) +
  labs(title   = "Total Projected Profit by Product Line (Jun-Dec 2005)",
       x       = "Total Projected Profit (USD Millions)",
       y       = NULL,
       caption = paste0("Company Total: ", dollar(round(overall_profit)))) +
  theme_minimal()

p_total
```

#### # Plot 3: Total Profit by Factory

This plot shows the total projected profit grouped by the factory that produces each line.

```{r kainen_plot_factory_profit, fig.height=5, fig.width=7}
profit_by_factory$factory_label <- paste0("Factory ", profit_by_factory$factory_id)

p_factory <- ggplot(profit_by_factory,
                    aes(x = factory_label,
                        y = total_profit / 1e6,
                        fill = factory_label)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  scale_fill_manual(values = c("Factory 0" = "#264653",
                               "Factory 1" = "#457B9D",
                               "Factory 2" = "#2A9D8F")) +
  scale_y_continuous(labels = dollar_format(suffix = "M", prefix = "$")) +
  labs(title = "Total Projected Profit by Factory (Jun-Dec 2005)",
       x     = NULL,
       y     = "Total Projected Profit (USD Millions)") +
  theme_minimal()

p_factory
```

#### # Plot 4 Monthly Projected Production by Product Line
```{r kainen_plot_forecast_monthly, fig.height=5, fig.width=10}
# Build a clean dataframe of the 7-month point forecasts
fc_monthly <- do.call(rbind, lapply(product_lines, function(pl) {
  fc <- fc_store[[pl]]
  data.frame(
    product_line = pl,
    month        = seq(as.Date("2005-06-01"), as.Date("2005-12-01"), by = "month"),
    month_label  = format(seq(as.Date("2005-06-01"), as.Date("2005-12-01"), by = "month"), "%b %Y"),
    forecast     = pmax(0, as.numeric(fc$mean)),
    lower80      = pmax(0, as.numeric(fc$lower[, 1])),
    upper80      = pmax(0, as.numeric(fc$upper[, 1]))
  )
}))

month_order <- format(seq(as.Date("2005-06-01"), as.Date("2005-12-01"), by = "month"), "%b %Y")
fc_monthly$month_label <- factor(fc_monthly$month_label, levels = month_order)

# forecasted units per product line per month
p_fc_monthly <- ggplot(fc_monthly,
                       aes(x = month_label, y = forecast, fill = product_line)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = PALETTE) +
  labs(title = "Forecasted Monthly Production by Product Line (Jun-Dec 2005)",
       x     = NULL,
       y     = "Forecasted Units Produced",
       fill  = "Product Line") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

p_fc_monthly
```
#### # Plot 5: Total Projected production by Product Line
```{r kainen_plot_forecast_annual, fig.height=5, fig.width=8}
fc_annual <- aggregate(forecast ~ product_line, data = fc_monthly, FUN = sum)
fc_annual <- fc_annual[order(-fc_annual$forecast), ]
fc_annual$product_line <- reorder(fc_annual$product_line, fc_annual$forecast)

p_fc_annual <- ggplot(fc_annual,
                      aes(x = forecast, y = product_line, fill = as.character(product_line))) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = round(forecast)), hjust = -0.2, size = 3.5) +
  scale_fill_manual(values = PALETTE) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title   = "Total Forecasted Production by Product Line (Jun-Dec 2005)",
       x       = "Total Forecasted Units",
       y       = NULL) +
  theme_minimal()

p_fc_annual
```
#### # Plot 6: Monthly Projected Profit by Product Line
```{r kainen_plot_profit_monthly_factory, fig.height=8, fig.width=12}
# Join factory info onto the monthly profit dataframe
fc_profit_factory <- merge(fc_profit, factory_map, by = "product_line", all.x = TRUE)
fc_profit_factory$factory_label <- paste0("Factory ", fc_profit_factory$factory_id)
fc_profit_factory$month_label   <- factor(fc_profit_factory$month_label, levels = month_order)

p_monthly_factory <- ggplot(fc_profit_factory,
                             aes(x = month_label, y = projected_profit / 1e3,
                                 fill = product_line)) +
  geom_col(width = 0.7) +
  facet_wrap(~ factory_label, ncol = 1) +
  scale_fill_manual(values = PALETTE) +
  scale_y_continuous(labels = dollar_format(suffix = "K", prefix = "$")) +
  labs(title   = "Projected Monthly Profit by Factory (Jun-Dec 2005)",
       x       = NULL,
       y       = "Projected Profit (USD Thousands)",
       fill    = "Product Line") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        strip.text  = element_text(face = "bold"))

p_monthly_factory
```
#### # Plot 7: Total Projected Profit by Factory and Product Line
```{r kainen_plot_profit_annual_factory, fig.height=5, fig.width=10}
# Summaries total Jun-Dec profit by factory and product line
profit_factory_pl <- aggregate(projected_profit ~ factory_label + product_line,
                                data = fc_profit_factory, FUN = sum)

p_annual_factory <- ggplot(profit_factory_pl,
                            aes(x = factory_label, y = projected_profit / 1e3,
                                fill = product_line)) +
  geom_col(width = 0.6) +
  scale_fill_manual(values = PALETTE) +
  scale_y_continuous(labels = dollar_format(suffix = "K", prefix = "$")) +
  labs(title   = "Total Projected Profit by Factory and Product Line (Jun-Dec 2005)",
       x       = NULL,
       y       = "Total Projected Profit (USD Thousands)",
       fill    = "Product Line") +
  theme_minimal()

p_annual_factory
```
#### # Plot 8: Total Projected Profit by Product Line
```{r kainen_plot_profit_annual_pl, fig.height=5, fig.width=10}
profit_pl_factory <- aggregate(projected_profit ~ product_line + factory_label,
                                data = fc_profit_factory, FUN = sum)
profit_pl_factory$product_line <- reorder(profit_pl_factory$product_line,
                                           profit_pl_factory$projected_profit)

p_annual_pl <- ggplot(profit_pl_factory,
                       aes(x = projected_profit / 1e3, y = product_line,
                           fill = factory_label)) +
  geom_col(width = 0.6) +
  scale_fill_manual(values = c("Factory 0" = "#264653",
                               "Factory 1" = "#457B9D",
                               "Factory 2" = "#2A9D8F")) +
  scale_x_continuous(labels = dollar_format(suffix = "K", prefix = "$")) +
  labs(title   = "Total Projected Profit by Product Line (Jun-Dec 2005)",
       x       = "Total Projected Profit (USD Thousands)",
       y       = NULL,
       fill    = "Factory") +
  theme_minimal()

p_annual_pl
```



