---
title: "Project_3_Production"
author: "Bea Sauve, Elton nichols, Kainen Osborne"
date: "2026-03-03"
output: html_document
---

# Project 3 Production Notebook
This notebook will entail all team members work including data refinement, forecasting productions, and statistic generations. Team member contributions will be divided into sections containing that specific persons work.
## ---------- Import Libraries ---------------------
```{r}
library(tidyverse)
library(readr)
library(lubridate)
```

## ---------- Production Data Refinement ---------------------
```{r production_refinement_Bea}
# Import data, convert weeks into date datatype
production_data <- read_csv('CSV/weekly_factory_production.csv',
                            col_types = cols(
                              week = col_date(format = "%Y-%m-%d")))
# Are there null vals
colSums((is.na(production_data)))

# Add month column
production_data$MONTH_PRODUCED <- format(production_data$week, "%Y-%m")

# Rename all columns to exclude spaces
production_data <- production_data %>%
  rename(
     "Classic_Cars" = `Classic Cars`,
     "Trucks_and_Buses" = `Trucks and Buses`,
     "Vintage_Cars" = `Vintage Cars` 
  )

# The outlier column indicates sections of weekly cost data that deviates from the norm. Create a flag column with 1 indicating that the row has outlieing data
production_data$outlier_flag <- if_else(production_data$outlier > 0, 1, 0)

# Calculate total production
production_data$TOTAL_UNITS <- production_data$Classic_Cars + production_data$Motorcycles + production_data$Planes + production_data$Ships + production_data$Trucks_and_Buses+production_data$Vintage_Cars

# Calculate item production costs
production_data$`AGGREGATE_CPU` <- production_data$weekly_cost / production_data$TOTAL_UNITS
```
##-----------Sales Data implementation------------
```{r}
# Import sales data, Convert N/A values back to North America, and format date
sales_data <- read_csv("CSV/sales_data_sample_cleaned.csv")
sales_data$TERRITORY[is.na(sales_data$TERRITORY)] <- "NA"
sales_data$ORDERDATE <- mdy(sales_data$ORDERDATE)

# Create a Month_Ordered Column containing the month and year combined
sales_data$MONTH_ORDERED <- format(sales_data$ORDERDATE, "%Y-%m")
```


## ---------------Regression Model for Productline Prices----------------
We would like to determine the actual cost per unit not based on aggregated data. 
A multiple linear regression model will assist in the isolation of each product cost under consideration that labor, cost of electricity, etc are consistent.
```{r product_regression_Bea}
regression_products <- lm(data = production_data, weekly_cost ~ Classic_Cars + Motorcycles +
                            Planes + Ships + Trucks_and_Buses + Vintage_Cars)
summary(regression_products)
```
Testing the models by factory
```{r production_by_factory_Bea}
F0_Production <- production_data %>% filter(factory_id == "0")
F1_Production <- production_data %>% filter(factory_id == "1")
F2_Production <- production_data %>% filter(factory_id == "2")

```
```{r factory_regression_bea}
f0_regression <- lm(data = F0_Production, weekly_cost ~ Classic_Cars + Ships)
summary(f0_regression)

f1_regression <- lm(data = F1_Production,weekly_cost ~  Motorcycles +Trucks_and_Buses)
summary(f1_regression)

f2_regression <- lm(data = F2_Production,weekly_cost ~ Planes+Vintage_Cars)
summary(f2_regression)

#include the factory for the regression
Master_Regression <- lm(weekly_cost ~ Classic_Cars + Motorcycles + Planes + Ships + 
                  Trucks_and_Buses + Vintage_Cars + factor(factory_id), 
                  data = production_data,
                  weights = if_else(production_data$outlier == 0, 1, 0))
summary(Master_Regression)
par(mfrow=c(2,2))
plot(Master_Regression)
par(mfrow=c(1,1))
```
## -----------Convert dataset into long format---------------------
```{r production_long_format_Bea}
# Crate a long format of production data
 
# Assign predicted coefficients of regression model to long format model
reg_coeffs <- coef(Master_Regression)
#Clean up the back ticks generated in production line names
names(reg_coeffs) <- gsub("`","", names(reg_coeffs))
# Define the base production intercept
base_intercept <- reg_coeffs["(Intercept)"]

# Implement regression calculated product cost per unit
weekly_productions_by_productline <- weekly_productions_by_productline %>%
  mutate(
    REGRESSION_CPU = ifelse(UNITS_PRODUCED > 0,
                            reg_coeffs[PRODUCTLINE],
                            0)
    )

# add intercepts based on factory id_ factory 0 is base ID
weekly_productions_by_productline <- weekly_productions_by_productline %>%
  mutate(
    REG_FACTORY_BASE_COST = base_intercept + case_when(
      factory_id == 1 ~ reg_coeffs["factor(factory_id)1"],
      factory_id == 2 ~ reg_coeffs["factor(factory_id)2"],
      TRUE            ~ 0 # This handles Factory 0
    )
  )
# add regression product costs
weekly_productions_by_productline$REG_TTL_PROD_COST_NO_BASE <- weekly_productions_by_productline$REGRESSION_CPU * weekly_productions_by_productline$UNITS_PRODUCED

weekly_productions_by_productline
```

## -------------Summarize weekly production into Monthly-------------------------
Condense the data set into a monthly data set instead of a weekly dataset. Implement regression model information into the dataset
```{r condensed_regression_data_Bea}
# condense the data by month and product line
monthly_productions_factory_X_productline <- weekly_productions_by_productline %>%
  group_by(Month, PRODUCTLINE, factory_id) %>%
  summarise(
    TOTAL_UNITS_PRODUCED = sum(UNITS_PRODUCED),
    .groups = "drop"
  ) %>% filter(TOTAL_UNITS_PRODUCED != 0)
# Implement regression calculated product cost per unit
monthly_productions_factory_X_productline <- monthly_productions_factory_X_productline %>%
  mutate(
    REGRESSION_CPU = ifelse(TOTAL_UNITS_PRODUCED > 0,
                            reg_coeffs[PRODUCTLINE],
                            0)
    )

# add intercepts based on factory id_ factory 0 is base ID
monthly_productions_factory_X_productline <-monthly_productions_factory_X_productline %>%
  mutate(
    REG_FACTORY_BASE_COST = base_intercept + case_when(
      factory_id == 1 ~ reg_coeffs["factor(factory_id)1"],
      factory_id == 2 ~ reg_coeffs["factor(factory_id)2"],
      TRUE            ~ 0 # This handles Factory 0
    )
  )
# add regression product costs without base
monthly_productions_factory_X_productline$REG_PROD_COST_NO_BASE <-monthly_productions_factory_X_productline$REGRESSION_CPU * monthly_productions_factory_X_productline$TOTAL_UNITS_PRODUCED
# add regression product costs with base
monthly_productions_factory_X_productline$AVG_PRODUCTION_COST_W_BASE <- monthly_productions_factory_X_productline$REGRESSION_CPU *
  monthly_productions_factory_X_productline$TOTAL_UNITS_PRODUCED + 
  monthly_productions_factory_X_productline$REG_FACTORY_BASE_COST

monthly_productions_factory_X_productline

```

#----------AGGREGATE MONTLHY DATA-----------
Note: Regression monthly calculations could be refined more
```{r Monthly_refinement_Bea}
#Calculate average price by product type- used for product monthly data
monthly_sales_by_productline <- sales_data %>%
  group_by(MONTH_ORDERED, PRODUCTLINE) %>%
  summarise(
    MIN_PRICE_SOLD = min(PRICEEACH, na.rm = TRUE),
      AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    MAX_PRICE_SOLD = max(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )
monthly_sales_by_productline

Final_SalesXProduction_Productline <- monthly_productions_factory_X_productline %>%
  full_join(monthly_sales_by_productline,
            by = c("Month" = "MONTH_ORDERED","PRODUCTLINE"))
# Calculate profits per product
Final_SalesXProduction_Productline$MIN_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$MIN_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
Final_SalesXProduction_Productline$AVG_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$AVG_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
Final_SalesXProduction_Productline$Max_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$MAX_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
            
# Add Average Revenue per product
Final_SalesXProduction_Productline$MIN_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$MIN_PRICE_SOLD
Final_SalesXProduction_Productline$AVG_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$AVG_PRICE_SOLD
Final_SalesXProduction_Productline$MAX_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$MAX_PRICE_SOLD
# Add average profit 
Final_SalesXProduction_Productline$MIN_TOTAL_PROFIT <- Final_SalesXProduction_Productline$MIN_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$AVG_TOTAL_PROFIT <- Final_SalesXProduction_Productline$AVG_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$MAX_TOTAL_PROFIT <- Final_SalesXProduction_Productline$MAX_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$ACTUAL_TOTAL_PROFIT <- Final_SalesXProduction_Productline$TOTAL_SALES - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE

Final_SalesXProduction_Productline
```

## ----------Export the Datasets for Visualizing-----------
```r
write_csv(Final_SalesXProduction_Productline, file.choose(new = TRUE))
write_csv(Final_SalesXProduction_Month, file.choose(new = TRUE))
```




## ---------- Bea's Production ---------------------


## ---------- Eltons's Production ---------------------


## ---------- Kainen's Production ---------------------