---
title: "Project_3_Production"
author: "Bea Sauve, Elton nichols, Kainen Osborne"
date: "2026-03-03"
output: html_document
---

# Project 3 Production Notebook
This notebook will entail all team members work including data refinement, forecasting productions, and statistic generations. Team member contributions will be divided into sections containing that specific persons work.
## ---------- Import Libraries ---------------------
```{r}
library(tidyverse)
library(readr)
library(lubridate)
```

## ---------- Production Data Refinement ---------------------
```{r production_refinement_Bea}
# Import data, convert weeks into date datatype
production_data <- read_csv('CSV/weekly_factory_production.csv',
                            col_types = cols(
                              week = col_date(format = "%Y-%m-%d")))
# Are there null vals
colSums((is.na(production_data)))

# Add month column
production_data$MONTH_PRODUCED <- format(production_data$week, "%Y-%m")

# Calculate total production
production_data$TOTAL_UNITS <- production_data$`Classic Cars` + production_data$Motorcycles + production_data$Planes + production_data$Ships + production_data$`Trucks and Buses`+production_data$`Vintage Cars`

# Calculate item production costs
production_data$`AGGREGATE_CPU` <- production_data$weekly_cost / production_data$TOTAL_UNITS
```
##-----------Sales Data implementation------------
```{r}
# Import sales data, Convert N/A values back to North America, and format date
sales_data <- read_csv("CSV/sales_data_sample_cleaned.csv")
sales_data$TERRITORY[is.na(sales_data$TERRITORY)] <- "NA"
sales_data$ORDERDATE <- mdy(sales_data$ORDERDATE)

# Create a Month_Ordered Column containing the month and year combined
sales_data$MONTH_ORDERED <- format(sales_data$ORDERDATE, "%Y-%m")
```


## ---------------Regression Model for Productline Prices----------------
We would like to determine the actual cost per unit not based on aggregated data. 
A multiple linear regression model will assist in the isolation of each product cost under consideration that labor, cost of electricity, etc are consistent.
```{r product_regression_Bea}
regression_products <- lm(data = production_data, weekly_cost ~ `Classic Cars` + Motorcycles +
                            Planes + Ships + `Trucks and Buses` + `Vintage Cars`)
summary(regression_products)
```
Testing the models by factory
```{r production_by_factory_Bea}
F0_Production <- production_data %>% filter(factory_id == "0")
F1_Production <- production_data %>% filter(factory_id == "1")
F2_Production <- production_data %>% filter(factory_id == "2")

```
```{r factory_regression_bea}
f0_regression <- lm(data = F0_Production, weekly_cost ~ `Classic Cars` + Ships)
summary(f0_regression)

f1_regression <- lm(data = F1_Production,weekly_cost ~  Motorcycles +`Trucks and Buses`)
summary(f1_regression)

f2_regression <- lm(data = F2_Production,weekly_cost ~ Planes+`Vintage Cars`)
summary(f2_regression)

#include the factory for the regression
Master_Regression <- lm(weekly_cost ~ `Classic Cars` + Motorcycles + Planes + Ships + 
                  `Trucks and Buses` + `Vintage Cars` + factor(factory_id) + outlier, 
                  data = production_data)
summary(Master_Regression)

```
## -----------Convert dataset into long format---------------------
```{r production_long_format_Bea}
# Crate a long format of production data
weekly_productions_by_productline <- production_data %>% 
  group_by(MONTH_PRODUCED) %>%
  pivot_longer(
    cols = c(`Classic Cars`,Motorcycles, Planes, Ships, `Trucks and Buses`, `Vintage Cars`),
    names_to = "PRODUCTLINE",
    values_to = "UNITS_PRODUCED"
  )
# Calculate product costs
weekly_productions_by_productline$AGG_PRODUCT_COSTS <- weekly_productions_by_productline$`AGGREGATE_CPU`* weekly_productions_by_productline$UNITS_PRODUCED
 
# Assign predicted coefficients of regression model to long format model
reg_coeffs <- coef(Master_Regression)
#Clean up the back ticks generated in production line names
names(reg_coeffs) <- gsub("`","", names(reg_coeffs))
# Define the base production intercept
base_intercept <- reg_coeffs["(Intercept)"]

# Implement regression calculated product cost per unit
weekly_productions_by_productline <- weekly_productions_by_productline %>%
  mutate(
    REGRESSION_CPU = ifelse(UNITS_PRODUCED > 0,
                            reg_coeffs[PRODUCTLINE],
                            0)
    )

# add intercepts based on factory id_ factory 0 is base ID
weekly_productions_by_productline <- weekly_productions_by_productline %>%
  mutate(
    REG_FACTORY_BASE_COST = base_intercept + case_when(
      factory_id == 1 ~ reg_coeffs["factor(factory_id)1"],
      factory_id == 2 ~ reg_coeffs["factor(factory_id)2"],
      TRUE            ~ 0 # This handles Factory 0
    )
  )
# add regression product costs
weekly_productions_by_productline$REG_TTL_PROD_COST_NO_BASE <-weekly_productions_by_productline$REGRESSION_CPU * weekly_productions_by_productline$UNITS_PRODUCED
```

## -------------Summarize weekly production into Monthly-------------------------
Condense the data set into a monthly data set instead of a weekly dataset. Implement regression model information into the dataset
```{r condensed_regression_data_Bea}
# condense the data by month and product line
monthly_productions_factory_X_productline <- weekly_productions_by_productline %>%
  group_by(MONTH_PRODUCED, PRODUCTLINE, factory_id) %>%
  summarise(
    TOTAL_UNITS_PRODUCED = sum(UNITS_PRODUCED),
    .groups = "drop"
  ) %>% filter(TOTAL_UNITS_PRODUCED != 0)
# Implement regression calculated product cost per unit
monthly_productions_factory_X_productline <- monthly_productions_factory_X_productline %>%
  mutate(
    REGRESSION_CPU = ifelse(TOTAL_UNITS_PRODUCED > 0,
                            reg_coeffs[PRODUCTLINE],
                            0)
    )

# add intercepts based on factory id_ factory 0 is base ID
monthly_productions_factory_X_productline <-monthly_productions_factory_X_productline %>%
  mutate(
    REG_FACTORY_BASE_COST = base_intercept + case_when(
      factory_id == 1 ~ reg_coeffs["factor(factory_id)1"],
      factory_id == 2 ~ reg_coeffs["factor(factory_id)2"],
      TRUE            ~ 0 # This handles Factory 0
    )
  )
# add regression product costs without base
monthly_productions_factory_X_productline$REG_PROD_COST_NO_BASE <-monthly_productions_factory_X_productline$REGRESSION_CPU * monthly_productions_factory_X_productline$TOTAL_UNITS_PRODUCED
# add regression product costs with base
monthly_productions_factory_X_productline$AVG_PRODUCTION_COST_W_BASE <- monthly_productions_factory_X_productline$REGRESSION_CPU *
  monthly_productions_factory_X_productline$TOTAL_UNITS_PRODUCED + 
  monthly_productions_factory_X_productline$REG_FACTORY_BASE_COST

```

#----------AGGREGATE MONTLHY DATA-----------
Note: Regression monthly calculations could be refined more
```{r Monthly_refinement_Bea}
# Aggregate the monthly data
#predicted weekly costs
production_data$PREDICTED_WEEKLY_COST <- predict(Master_Regression, newdata = production_data)

Summary_Monthly_Data <- production_data %>%
  group_by(MONTH_PRODUCED) %>%
  summarise(
    MONTHLY_PRODUCTION  = sum(TOTAL_UNITS),
    ACTUAL_MONTHLY_COST = sum(weekly_cost),
    PRED_MONTHLY_COST   = sum(PREDICTED_WEEKLY_COST),
    #Actual Cost Per Unit
    ACTUAL_AVG_CPU      = ACTUAL_MONTHLY_COST / MONTHLY_PRODUCTION,
    # Predicted Cost Per Unit
    REGRESSION_AVG_CPU  = PRED_MONTHLY_COST / MONTHLY_PRODUCTION,
    CPU_VARIANCE        = ACTUAL_AVG_CPU - REGRESSION_AVG_CPU
  )

#Calculate average price by month - used for monthly data
Monthly_Sales_Data <- sales_data %>%
  group_by(MONTH_ORDERED) %>%
  summarise(
    MIN_PRICE_SOLD = min(PRICEEACH, na.rm = TRUE),
    AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    MAX_PRICE_SOLD = max(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )

# FUll join of sales and production
Final_SalesXProduction_Month <- Summary_Monthly_Data %>%
  full_join(Monthly_Sales_Data,
            by = c("MONTH_PRODUCED" = "MONTH_ORDERED"))
#add Revenue
Final_SalesXProduction_Month$MIN_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$MIN_PRICE_SOLD
Final_SalesXProduction_Month$AVG_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$AVG_PRICE_SOLD
Final_SalesXProduction_Month$MAX_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$MAX_PRICE_SOLD

#add profit
Final_SalesXProduction_Month$MIN_PROFIT <- Final_SalesXProduction_Month$MIN_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$AVG_PROFIT <- Final_SalesXProduction_Month$AVG_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$MAX_PROFIT <- Final_SalesXProduction_Month$MAX_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$ACTUAL_PROFIT <- Final_SalesXProduction_Month$TOTAL_SALES - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
```

## ---------- Combine Sales and Production by Product ---------------------
```{r Combine_sales_production_by_product_Bea}
#Calculate average price by product type- used for product monthly data
monthly_sales_by_productline <- sales_data %>%
  group_by(MONTH_ORDERED, PRODUCTLINE) %>%
  summarise(
    MIN_PRICE_SOLD = min(PRICEEACH, na.rm = TRUE),
      AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    MAX_PRICE_SOLD = max(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )

# Implement a Full join of the condensed production data and sales data
Final_SalesXProduction_Productline <- monthly_productions_factory_X_productline %>%
  full_join(monthly_sales_by_productline,
            by = c("MONTH_PRODUCED" = "MONTH_ORDERED","PRODUCTLINE"))
            
# Add Average Revenue per product
Final_SalesXProduction_Productline$MIN_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$MIN_PRICE_SOLD
Final_SalesXProduction_Productline$AVG_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$AVG_PRICE_SOLD
Final_SalesXProduction_Productline$MAX_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$MAX_PRICE_SOLD
# Add average profit 
Final_SalesXProduction_Productline$MIN_PROFIT <- Final_SalesXProduction_Productline$MIN_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$AVG_PROFIT <- Final_SalesXProduction_Productline$AVG_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$MAX_PROFIT <- Final_SalesXProduction_Productline$MAX_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$ACTUAL_PROFIT <- Final_SalesXProduction_Productline$TOTAL_SALES - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE

```
## ----------Export the Datasets for Visualizing-----------
```r
write_csv(Final_SalesXProduction_Productline, file.choose(new = TRUE))
write_csv(Final_SalesXProduction_Month, file.choose(new = TRUE))
```




## ---------- Bea's Production ---------------------


## ---------- Eltons's Production ---------------------

## ---------- Kainen's Production ---------------------

```{r kainen_setup}
suppressPackageStartupMessages({
  library(forecast)
  library(tseries)
  library(scales)
  library(gridExtra)
})

# Add date column to joined product-line dataframe
Final_SalesXProduction_Productline$date <- as.Date(
  paste0(Final_SalesXProduction_Productline$MONTH_PRODUCED, "-01")
)

# Color palette used across all of Kainen's plots
# Codes are taken from online.
PALETTE <- c(
  "Classic Cars"     = "#E63946",
  "Motorcycles"      = "#457B9D",
  "Planes"           = "#2A9D8F",
  "Ships"            = "#E9C46A",
  "Trucks and Buses" = "#6A0572",
  "Vintage Cars"     = "#264653"
)

# Product lines with tracked production
product_lines <- c("Classic Cars", "Motorcycles", "Planes",
                   "Ships", "Trucks and Buses", "Vintage Cars")

forecast_horizon <- 7
```

### | Monthly Time Series Objects

This converts the cleaned product-line dataframe into one time series object per
product line.

```{r kainen_build_ts}
ts_list <- list()

for (pl in product_lines) {
  pl_data <- Final_SalesXProduction_Productline %>%
    filter(PRODUCTLINE == pl, !is.na(TOTAL_UNITS_PRODUCED)) %>%
    arrange(date)

  ts_list[[pl]] <- ts(
    pl_data$TOTAL_UNITS_PRODUCED,
    start     = c(year(min(pl_data$date)), month(min(pl_data$date))),
    frequency = 12
  )
}

# Print coverage for each series
for (pl in product_lines) {
  cat(sprintf("%s:  start = %s  end = %s  n = %d obs\n",
              pl,
              paste(start(ts_list[[pl]]), collapse = "-"),
              paste(end(ts_list[[pl]]),   collapse = "-"),
              length(ts_list[[pl]])))
}
```



### | Overall Production of each Product Line.

```{r kainen_plot_series, fig.height=10, fig.width=12}
series_plots <- list()

for (pl in product_lines) {
  series_plots[[pl]] <- autoplot(ts_list[[pl]]) +
    ggtitle(pl) +
    xlab("Year") +
    ylab("Units Produced") +
    theme_minimal()
}

do.call(grid.arrange, c(series_plots, list(ncol = 2,
        top  = "Monthly Units Produced by Product Line (Jan 2003 - May 2005)")))
```


### | Stationarity and Seasonality Checks


```{r kainen_ndiffs}
# Checking for difference
for (pl in product_lines) {
  d <- ndiffs(ts_list[[pl]])
  cat(sprintf("%-20s  ndiffs = %d\n", pl, d))
}
```

### | Augmented Dickey-Fuller test

The ADF test checks whether the series has a unit root (i.e. is
non-stationary).

If p < 0.05 we have stationarity.
If p >= 0.05 there is evidence of non-stationarity.

```{r kainen_adf}
for (pl in product_lines) {
  result <- adf.test(ts_list[[pl]])
  cat(sprintf("%-20s  ADF p-value = %.4f  (%s)\n",
              pl,
              result$p.value,
              ifelse(result$p.value < 0.05, "stationary", "non-stationary")))
}
```

### | Seasonality detection via STL

```{r kainen_seasonality}
for (pl in product_lines) {
  ts_obj  <- ts_list[[pl]]
  stl_fit <- stl(ts_obj, s.window = "periodic")
  sr <- var(stl_fit$time.series[, "seasonal"]) / var(ts_obj)
  cat(sprintf("%-20s  seasonal ratio = %.3f  (%s)\n",
              pl,
              sr,
              ifelse(sr > 0.10, "seasonality detected", "weak or no seasonality")))
}
```

### | Comparing All Four Model Types

```{r kainen_fit_models}
fit_store    <- list()
model_labels <- list()
aicc_table   <- data.frame()

for (pl in product_lines) {
  ts_obj <- ts_list[[pl]]
  n      <- length(ts_obj)

  # -- ARIMA --
  fit_arima <- tryCatch(
    auto.arima(ts_obj, stepwise = FALSE, approximation = FALSE, ic = "aicc"),
    error = function(e) NULL
  )

  # -- ETS --
  fit_ets <- tryCatch(
    ets(ts_obj, ic = "aicc"),
    error = function(e) NULL
  )

  # -- STL (ETS on seasonally adjusted component) --
  fit_stl <- tryCatch(
    stlf(ts_obj, h = forecast_horizon),
    error = function(e) NULL
  )

  # -- Linear regression with trend + season --
  fit_lm <- tryCatch(
    tslm(ts_obj ~ trend + season),
    error = function(e) NULL
  )

  # Collect AICc values (STL and LM use AIC as proxy)
  aicc_arima <- if (!is.null(fit_arima)) fit_arima$aicc   else NA
  aicc_ets   <- if (!is.null(fit_ets))   fit_ets$aicc     else NA
  aicc_stl   <- if (!is.null(fit_stl))   AIC(fit_stl$model) else NA
  aicc_lm    <- if (!is.null(fit_lm))    AIC(fit_lm)      else NA

  aicc_table <- rbind(aicc_table, data.frame(
    product_line = pl,
    ARIMA        = round(aicc_arima, 1),
    ETS          = round(aicc_ets,   1),
    STL          = round(aicc_stl,   1),
    LM           = round(aicc_lm,    1)
  ))

  # Select best model by lowest AICc
  candidates <- c(ARIMA = aicc_arima, ETS = aicc_ets, STL = aicc_stl, LM = aicc_lm)
  best_name  <- names(which.min(candidates))

  best_fit <- switch(best_name,
    ARIMA = fit_arima,
    ETS   = fit_ets,
    STL   = fit_stl$model,
    LM    = fit_lm
  )

  fit_store[[pl]]    <- list(arima = fit_arima, ets = fit_ets,
                             stl   = fit_stl,   lm  = fit_lm,
                             best  = best_fit,  best_name = best_name)
  model_labels[[pl]] <- best_name

  cat(sprintf("\n%s -- selected: %s  (AICc: ARIMA=%.1f  ETS=%.1f  STL=%.1f  LM=%.1f)\n",
              pl, best_name,
              aicc_arima, aicc_ets, aicc_stl, aicc_lm))
}
```

### | AICc Summary Table

```{r kainen_aicc_table}
print(aicc_table, row.names = FALSE)
```

### | Accuracy Comparison

- **RMSE** -- root mean squared error. Penalizes large errors more heavily.
- **MAE** -- mean absolute error. Average absolute deviation from true values.
- **MAPE** -- mean absolute percentage error. Error as a percentage of true value,
  easier to interpret across series with different scales.
- **MASE** -- compares the model against the seasonal naive baseline. Values below
  1.0 mean the model beats simply repeating last year's value.
- **ACF1** -- autocorrelation of residuals at lag 1. Values near zero mean the
  model has captured the serial structure and left behind noise.

```{r kainen_accuracy, warning=FALSE}
for (pl in product_lines) {
  ts_obj <- ts_list[[pl]]
  fits   <- fit_store[[pl]]

  fc_arima <- if (!is.null(fits$arima)) forecast(fits$arima, h = forecast_horizon) else NULL
  fc_ets   <- if (!is.null(fits$ets))   forecast(fits$ets,   h = forecast_horizon) else NULL
  fc_stl   <- fits$stl
  fc_lm    <- if (!is.null(fits$lm))    forecast(fits$lm,    h = forecast_horizon) else NULL

  rows <- list()
  if (!is.null(fc_arima)) rows[["ARIMA"]] <- accuracy(fc_arima)[1, ]
  if (!is.null(fc_ets))   rows[["ETS"]]   <- accuracy(fc_ets)[1, ]
  if (!is.null(fc_stl))   rows[["STL"]]   <- accuracy(fc_stl)[1, ]
  if (!is.null(fc_lm))    rows[["LM"]]    <- accuracy(fc_lm)[1, ]

  if (length(rows) > 0) {
    tbl <- round(do.call(rbind, rows), 2)
    cat(sprintf("\n--- %s (selected: %s) ---\n", pl, model_labels[[pl]]))
    print(tbl)
  }
}
```

### | Model Diagnostics

```{r kainen_diagnostics, fig.height=5, fig.width=10}
for (pl in product_lines) {
  fits      <- fit_store[[pl]]
  best_name <- model_labels[[pl]]
  best_fit  <- fits$best

  cat(sprintf("\n=== %s -- diagnostics for selected model: %s ===\n", pl, best_name))
  checkresiduals(best_fit)
}
```



### | Forecasts for June to December 2005

This step generates a 7-month forecast for each product line using the selected
model. Each plot shows:

- **Solid black line** : the historical production series (Jan 2003 - May 2005)
- **Solid Blue line** : the point forecast for Jun - Dec 2005
- **Dark shaded band** : 80% prediction interval
- **Light shaded band** : 95% prediction interval

```{r kainen_forecasts, fig.height=5, fig.width=10}
forecast_dates <- seq(as.Date("2005-06-01"), as.Date("2005-12-01"), by = "month")

for (pl in product_lines) {
  fits      <- fit_store[[pl]]
  best_name <- model_labels[[pl]]

  fc <- fits$stl

  fc_store[[pl]] <- fc

  print(
    autoplot(fc) +
      xlim(2003, 2006) +
      ggtitle(paste(pl, "-- forecast Jun-Dec 2005 (", best_name, ")")) +
      xlab("Month") +
      ylab("Units Produced") +
      theme_minimal()
  )
}
```


### | Average Profit Per Unit


```{r kainen_ppu}
profit_per_unit <- Final_SalesXProduction_Productline %>%
  filter(!is.na(ACTUAL_PROFIT),
         !is.na(TOTAL_UNITS_PRODUCED),
         TOTAL_UNITS_PRODUCED > 0) %>%
  mutate(ppu = ACTUAL_PROFIT / TOTAL_UNITS_PRODUCED) %>%
  group_by(PRODUCTLINE) %>%
  summarise(avg_profit_per_unit = mean(ppu, na.rm = TRUE),
            n_months = n(),
            .groups = "drop") %>%
  rename(product_line = PRODUCTLINE) %>%
  arrange(desc(avg_profit_per_unit))

print(profit_per_unit, digits = 2)
```

### | Profit Projections for June to December 2005

```{r kainen_profit_calc}
# Build a dataframe of point forecasts for Jun-Dec 2005
forecast_months <- seq(as.Date("2005-06-01"), as.Date("2005-12-01"), by = "month")

fc_rows <- list()
for (pl in product_lines) {
  fc     <- fc_store[[pl]]
  fc_rows[[pl]] <- data.frame(
    product_line = pl,
    month        = forecast_months,
    forecast     = pmax(0, as.numeric(fc$mean)),
    lower80      = pmax(0, as.numeric(fc$lower[, 1])),
    upper80      = pmax(0, as.numeric(fc$upper[, 1]))
  )
}

fc_df <- do.call(rbind, fc_rows)

# Join profit per unit and calculate projected profit
fc_profit <- merge(fc_df,
                   profit_per_unit[, c("product_line", "avg_profit_per_unit")],
                   by = "product_line")

fc_profit$projected_profit <- fc_profit$forecast   * fc_profit$avg_profit_per_unit
fc_profit$profit_lo80      <- fc_profit$lower80    * fc_profit$avg_profit_per_unit
fc_profit$profit_hi80      <- fc_profit$upper80    * fc_profit$avg_profit_per_unit
fc_profit$month_label      <- format(fc_profit$month, "%b %Y")

# Summary by product line
profit_by_pl <- aggregate(
  cbind(total_units = forecast,
        total_profit = projected_profit,
        lower_80 = profit_lo80,
        upper_80 = profit_hi80) ~ product_line,
  data = fc_profit, FUN = sum
)
profit_by_pl <- profit_by_pl[order(-profit_by_pl$total_profit), ]

# Factory map
factory_map <- unique(Final_SalesXProduction_Productline[
  !is.na(Final_SalesXProduction_Productline$factory_id),
  c("PRODUCTLINE", "factory_id")
])
names(factory_map)[1] <- "product_line"
factory_map <- factory_map[!duplicated(factory_map$product_line), ]

profit_by_factory <- merge(fc_profit, factory_map, by = "product_line", all.x = TRUE)
profit_by_factory <- aggregate(
  cbind(total_profit = projected_profit,
        lower_80 = profit_lo80,
        upper_80 = profit_hi80) ~ factory_id,
  data = profit_by_factory[!is.na(profit_by_factory$factory_id), ],
  FUN = sum
)

overall_profit <- sum(profit_by_pl$total_profit, na.rm = TRUE)

cat("\n=== Projected Profit by Product Line (Jun-Dec 2005) ===\n")
profit_by_pl$total_profit_fmt <- dollar(round(profit_by_pl$total_profit))
profit_by_pl$lower_80_fmt     <- dollar(round(profit_by_pl$lower_80))
profit_by_pl$upper_80_fmt     <- dollar(round(profit_by_pl$upper_80))
print(profit_by_pl[, c("product_line", "total_units",
                        "total_profit_fmt", "lower_80_fmt", "upper_80_fmt")],
      row.names = FALSE)

cat(sprintf("\nCompany Total (excl. Trains): %s\n", dollar(round(overall_profit))))

cat("\n=== Projected Profit by Factory (Jun-Dec 2005) ===\n")
print(profit_by_factory, row.names = FALSE)
```

### | Profit Projection Plots

#### # Plot 1: Monthly Stacked Profit

This plot shows projected profit for each month from June to December 2005,
broken down by product line.

```{r kainen_plot_monthly_profit, fig.height=5, fig.width=10}
month_order <- format(forecast_months, "%b %Y")
fc_profit$month_label <- factor(fc_profit$month_label, levels = month_order)

p_monthly <- ggplot(fc_profit,
                    aes(x = month_label,
                        y = projected_profit / 1e6,
                        fill = product_line)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = PALETTE) +
  scale_y_continuous(labels = dollar_format(suffix = "M", prefix = "$")) +
  labs(title   = "Projected Monthly Profit by Product Line (Jun-Dec 2005)",
       x       = NULL,
       y       = "Projected Profit (USD Millions)",
       fill    = "Product Line",
       caption = "Point forecast x average historical profit per unit") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

p_monthly
```

#### # Plot 2: Total Profit by Product Line

This bar chart ranks each product line by its total projected profit across the
full June to December 2005 period.

```{r kainen_plot_total_profit, fig.height=5, fig.width=9}
profit_by_pl$product_line <- reorder(profit_by_pl$product_line,
                                     profit_by_pl$total_profit)

p_total <- ggplot(profit_by_pl,
                  aes(x = total_profit / 1e6,
                      y = product_line,
                      fill = as.character(product_line))) +
  geom_col(width = 0.6, show.legend = FALSE) +
  scale_fill_manual(values = PALETTE) +
  scale_x_continuous(labels = dollar_format(suffix = "M", prefix = "$")) +
  labs(title   = "Total Projected Profit by Product Line (Jun-Dec 2005)",
       x       = "Total Projected Profit (USD Millions)",
       y       = NULL,
  theme_minimal())

p_total
```

#### # Plot 3: Total Profit by Factory

This plot shows the total projected profit grouped by the factory that produces each line.

```{r kainen_plot_factory_profit, fig.height=5, fig.width=7}
profit_by_factory$factory_label <- paste0("Factory ", profit_by_factory$factory_id)

p_factory <- ggplot(profit_by_factory,
                    aes(x = factory_label,
                        y = total_profit / 1e6,
                        fill = factory_label)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  scale_fill_manual(values = c("Factory 0" = "#264653",
                               "Factory 1" = "#457B9D",
                               "Factory 2" = "#2A9D8F")) +
  scale_y_continuous(labels = dollar_format(suffix = "M", prefix = "$")) +
  labs(title = "Total Projected Profit by Factory (Jun-Dec 2005)",
       x     = NULL,
       y     = "Total Projected Profit (USD Millions)") +
  theme_minimal()

p_factory
```

#### # Plot 4: ACF by Product Line

The ACF (autocorrelation function) plot shows how correlated each month's
production value is with values from previous months (lags).

```{r kainen_plot_acf, fig.height=10, fig.width=12}
acf_plots <- list()

for (pl in product_lines) {
  ts_obj   <- ts_list[[pl]]
  acf_data <- acf(ts_obj, plot = FALSE, lag.max = 20)
  df_acf   <- data.frame(lag = as.numeric(acf_data$lag),
                         acf = as.numeric(acf_data$acf))
  ci <- qnorm(0.975) / sqrt(length(ts_obj))

  acf_plots[[pl]] <- ggplot(df_acf, aes(x = lag, y = acf)) +
    geom_hline(yintercept = 0, color = "grey50") +
    geom_hline(yintercept = c(-ci, ci), linetype = "dashed", color = "steelblue") +
    geom_segment(aes(xend = lag, y = 0, yend = acf), color = PALETTE[pl]) +
    geom_point(color = PALETTE[pl], size = 1.5) +
    ylim(-1, 1) +
    labs(title = pl, x = "Lag (months)", y = "ACF") +
    theme_minimal()
}

g_acf <- do.call(grid.arrange,
                 c(acf_plots, list(ncol = 2, top = "ACF by Product Line")))

```

#### # Plot 5: Model Residuals

The residual plot shows what is left over after the selected model has done its
job. If the model captured all the structure in the series, the residuals should
look like random noise with no seasonality or trends.

```{r kainen_plot_residuals, fig.height=10, fig.width=12}
resid_plots <- list()

for (pl in product_lines) {
  fits      <- fit_store[[pl]]
  best_name <- model_labels[[pl]]

  res <- residuals(fits$best)
  df_res <- data.frame(t = as.numeric(time(res)), resid = as.numeric(res))

  resid_plots[[pl]] <- ggplot(df_res, aes(x = t, y = resid)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_line(color = PALETTE[pl]) +
    geom_point(color = PALETTE[pl], size = 1.5) +
    labs(title    = pl,
         subtitle = paste("Model:", best_name),
         x        = "Time",
         y        = "Residual") +
    theme_minimal()
}

g_resid <- do.call(grid.arrange,
                   c(resid_plots, list(ncol = 2, top = "Model Residuals by Product Line")))

```

