---
title: "Project_3_Production"
author: "Bea Sauve, Elton nichols, Kainen Osborne"
date: "2026-03-03"
output: html_document
---

# Project 3 Production Notebook

This notebook will entail all team members work including data refinement, forecasting productions, and statistic generations. Team member contributions will be divided into sections containing that specific persons work. \## ---------- Import Libraries ---------------------

```{r}
library(tidyverse)
library(readr)       # Both readr and lubridate should be included already when installing tidyverse
library(lubridate)
library(forecast)
library(tseries)
```

## ---------- Production Data Refinement ---------------------

```{r production_refinement_Bea}
# Import data, convert weeks into date datatype
production_data <- read_csv('CSV/weekly_factory_production.csv',
                            col_types = cols(
                              week = col_date(format = "%Y-%m-%d")))
# Are there null vals
colSums((is.na(production_data)))

# Add month column
production_data$MONTH_PRODUCED <- format(production_data$week, "%Y-%m")

# Rename all columns to exclude spaces
production_data <- production_data %>%
  rename(
     "Classic_Cars" = `Classic Cars`,
     "Trucks_and_Buses" = `Trucks and Buses`,
     "Vintage_Cars" = `Vintage Cars` 
  )


# The outlier column indicates sections of weekly cost data that deviates from the norm. Create a flag column with 1 indicating that the row has outlieing data
production_data$outlier_flag <- if_else(production_data$outlier > 0, 1, 0)

# Calculate total production
production_data$TOTAL_UNITS <- production_data$Classic_Cars + production_data$Motorcycles + production_data$Planes + production_data$Ships + production_data$Trucks_and_Buses+production_data$Vintage_Cars

# Calculate item production costs
production_data$`AGGREGATE_CPU` <- production_data$weekly_cost / production_data$TOTAL_UNITS
```

##-----------Sales Data implementation------------

```{r}
# Import sales data, Convert N/A values back to North America, and format date
sales_data <- read_csv("CSV/sales_data_sample_cleaned.csv")
sales_data$TERRITORY[is.na(sales_data$TERRITORY)] <- "NA"
sales_data$ORDERDATE <- mdy(sales_data$ORDERDATE)

# Create a Month_Ordered Column containing the month and year combined
sales_data$MONTH_ORDERED <- format(sales_data$ORDERDATE, "%Y-%m")

sales_data <- sales_data %>%
  mutate(PRODUCTLINE = gsub(" ", "_", PRODUCTLINE))

```

## ---------------Regression Model for Productline Prices----------------

We would like to determine the actual cost per unit not based on aggregated data. A multiple linear regression model will assist in the isolation of each product cost under consideration that labor, cost of electricity, etc are consistent.

```{r product_regression_Bea}
regression_products <- lm(data = production_data, weekly_cost ~ Classic_Cars + Motorcycles +
                            Planes + Ships + Trucks_and_Buses + Vintage_Cars)
summary(regression_products)
```

Testing the models by factory

```{r production_by_factory_Bea}
F0_Production <- production_data %>% filter(factory_id == "0")
F1_Production <- production_data %>% filter(factory_id == "1")
F2_Production <- production_data %>% filter(factory_id == "2")

```

```{r factory_regression_bea}
f0_regression <- lm(data = F0_Production, weekly_cost ~ Classic_Cars + Ships)
summary(f0_regression)

f1_regression <- lm(data = F1_Production,weekly_cost ~  Motorcycles +Trucks_and_Buses)
summary(f1_regression)

f2_regression <- lm(data = F2_Production,weekly_cost ~ Planes+Vintage_Cars)
summary(f2_regression)

#include the factory for the regression
Master_Regression <- lm(weekly_cost ~ Classic_Cars + Motorcycles + Planes + Ships + 
                  Trucks_and_Buses + Vintage_Cars + factor(factory_id), 
                  data = production_data,
                  weights = if_else(production_data$outlier == 0, 1, 0))
summary(Master_Regression)
par(mfrow=c(2,2))
plot(Master_Regression)
par(mfrow=c(1,1))
```

## -----------Convert dataset into long format---------------------

```{r production_long_format_Bea}
# Crate a long format of production data
weekly_productions_by_productline <- production_data %>% 
  group_by(MONTH_PRODUCED) %>%
  pivot_longer(
    cols = c(Classic_Cars,Motorcycles, Planes, Ships, Trucks_and_Buses, Vintage_Cars),
    names_to = "PRODUCTLINE",
    values_to = "UNITS_PRODUCED"
  )
# Calculate product costs
weekly_productions_by_productline$AGG_PRODUCT_COSTS <- weekly_productions_by_productline$`AGGREGATE_CPU`* weekly_productions_by_productline$UNITS_PRODUCED
 
# Assign predicted coefficients of regression model to long format model
reg_coeffs <- coef(Master_Regression)
#Clean up the back ticks generated in production line names
names(reg_coeffs) <- gsub("`","", names(reg_coeffs))
# Define the base production intercept
base_intercept <- reg_coeffs["(Intercept)"]

# Implement regression calculated product cost per unit
weekly_productions_by_productline <- weekly_productions_by_productline %>%
  mutate(
    REGRESSION_CPU = ifelse(TOTAL_UNITS > 0,
                            reg_coeffs[PRODUCTLINE],
                            0)
    )

# add intercepts based on factory id_ factory 0 is base ID
weekly_productions_by_productline <- weekly_productions_by_productline %>%
  mutate(
    REG_FACTORY_BASE_COST = base_intercept + case_when(
      factory_id == 1 ~ reg_coeffs["factor(factory_id)1"],
      factory_id == 2 ~ reg_coeffs["factor(factory_id)2"],
      TRUE            ~ 0 # This handles Factory 0
    )
  )
# add regression product costs
weekly_productions_by_productline$REG_TTL_PROD_COST_NO_BASE <- weekly_productions_by_productline$REGRESSION_CPU * weekly_productions_by_productline$UNITS_PRODUCED

weekly_productions_by_productline
```

## -------------Summarize weekly production into Monthly-------------------------

Condense the data set into a monthly data set instead of a weekly dataset. Implement regression model information into the dataset

```{r condensed_regression_data_Bea}
# condense the data by month and product line
monthly_productions_factory_X_productline <- weekly_productions_by_productline %>%
  group_by(MONTH_PRODUCED, PRODUCTLINE, factory_id) %>%
  summarise(
    TOTAL_UNITS_PRODUCED = sum(UNITS_PRODUCED),
    .groups = "drop"
  ) %>% filter(TOTAL_UNITS_PRODUCED != 0)
# Implement regression calculated product cost per unit
monthly_productions_factory_X_productline <- monthly_productions_factory_X_productline %>%
  mutate(
    REGRESSION_CPU = ifelse(TOTAL_UNITS_PRODUCED > 0,
                            reg_coeffs[PRODUCTLINE],
                            0)
    )

# add intercepts based on factory id_ factory 0 is base ID
monthly_productions_factory_X_productline <-monthly_productions_factory_X_productline %>%
  mutate(
    REG_FACTORY_BASE_COST = base_intercept + case_when(
      factory_id == 1 ~ reg_coeffs["factor(factory_id)1"],
      factory_id == 2 ~ reg_coeffs["factor(factory_id)2"],
      TRUE            ~ 0 # This handles Factory 0
    )
  )
# add regression product costs without base
monthly_productions_factory_X_productline$REG_PROD_COST_NO_BASE <-monthly_productions_factory_X_productline$REGRESSION_CPU * monthly_productions_factory_X_productline$TOTAL_UNITS_PRODUCED
# add regression product costs with base
monthly_productions_factory_X_productline$AVG_PRODUCTION_COST_W_BASE <- monthly_productions_factory_X_productline$REGRESSION_CPU *
  monthly_productions_factory_X_productline$TOTAL_UNITS_PRODUCED + 
  monthly_productions_factory_X_productline$REG_FACTORY_BASE_COST

monthly_productions_factory_X_productline

```

#----------AGGREGATE MONTLHY DATA----------- Note: Regression monthly calculations could be refined more

```{r Monthly_refinement_Bea}
#Calculate average price by product type- used for product monthly data
monthly_sales_by_productline <- sales_data %>%
  group_by(MONTH_ORDERED, PRODUCTLINE) %>%
  summarise(
    MIN_PRICE_SOLD = min(PRICEEACH, na.rm = TRUE),
      AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    MAX_PRICE_SOLD = max(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )
monthly_sales_by_productline

Final_SalesXProduction_Productline <- monthly_productions_factory_X_productline %>%
  full_join(monthly_sales_by_productline,
            by = c("MONTH_PRODUCED" = "MONTH_ORDERED","PRODUCTLINE"))
# Calculate profits per product
Final_SalesXProduction_Productline$MIN_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$MIN_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
Final_SalesXProduction_Productline$AVG_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$AVG_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
Final_SalesXProduction_Productline$Max_PRODUCT_PROFIT <- Final_SalesXProduction_Productline$MAX_PRICE_SOLD - Final_SalesXProduction_Productline$REGRESSION_CPU
            
# Add Average Revenue per product
Final_SalesXProduction_Productline$MIN_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$MIN_PRICE_SOLD
Final_SalesXProduction_Productline$AVG_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$AVG_PRICE_SOLD
Final_SalesXProduction_Productline$MAX_TOTAL_REVENUE <- Final_SalesXProduction_Productline$TOTAL_SHIPPED * Final_SalesXProduction_Productline$MAX_PRICE_SOLD
# Add average profit 
Final_SalesXProduction_Productline$MIN_TOTAL_PROFIT <- Final_SalesXProduction_Productline$MIN_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$AVG_TOTAL_PROFIT <- Final_SalesXProduction_Productline$AVG_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$MAX_TOTAL_PROFIT <- Final_SalesXProduction_Productline$MAX_TOTAL_REVENUE - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE
Final_SalesXProduction_Productline$ACTUAL_TOTAL_PROFIT <- Final_SalesXProduction_Productline$TOTAL_SALES - Final_SalesXProduction_Productline$AVG_PRODUCTION_COST_W_BASE

Final_SalesXProduction_Productline
```

Aggregate monthly sales no product line

```{r}
#predicted weekly costs
production_data$PREDICTED_WEEKLY_COST <- predict(Master_Regression, newdata = production_data)

Summary_Monthly_Data <- production_data %>%
  group_by(MONTH_PRODUCED) %>%
  summarise(
    MONTHLY_PRODUCTION  = sum(TOTAL_UNITS),
    ACTUAL_MONTHLY_COST = sum(weekly_cost),
    PRED_MONTHLY_COST   = sum(PREDICTED_WEEKLY_COST),
    #Actual Cost Per Unit
    ACTUAL_AVG_CPU      = ACTUAL_MONTHLY_COST / MONTHLY_PRODUCTION,
    # Predicted Cost Per Unit
    REGRESSION_AVG_CPU  = PRED_MONTHLY_COST / MONTHLY_PRODUCTION,
    CPU_VARIANCE        = ACTUAL_AVG_CPU - REGRESSION_AVG_CPU
  )
#Calculate average price by month - used for monthly data
Monthly_Sales_Data <- sales_data %>%
  group_by(MONTH_ORDERED) %>%
  summarise(
    MIN_PRICE_SOLD = min(PRICEEACH, na.rm = TRUE),
    AVG_PRICE_SOLD = mean(PRICEEACH, na.rm = TRUE),
    MAX_PRICE_SOLD = max(PRICEEACH, na.rm = TRUE),
    AVG_SHIPPED_PER_ORDER = mean(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SHIPPED = sum(QUANTITYORDERED, na.rm = TRUE),
    TOTAL_SALES = sum(SALES)
  )
# FUll join of sales and production
Final_SalesXProduction_Month <- Summary_Monthly_Data %>%
  full_join(Monthly_Sales_Data,
            by = c("MONTH_PRODUCED" = "MONTH_ORDERED"))
# Ensure Dates are date data types
Final_SalesXProduction_Productline$Month <- ym(Final_SalesXProduction_Productline$MONTH_PRODUCED)

# Calculate profits per product
Final_SalesXProduction_Month$MIN_PRODUCT_PROFIT <- Final_SalesXProduction_Month$MIN_PRICE_SOLD - Final_SalesXProduction_Month$REGRESSION_AVG_CPU
Final_SalesXProduction_Month$AVG_PRODUCT_PROFIT <- Final_SalesXProduction_Month$AVG_PRICE_SOLD - Final_SalesXProduction_Month$REGRESSION_AVG_CPU
Final_SalesXProduction_Month$Max_PRODUCT_PROFIT <- Final_SalesXProduction_Month$MAX_PRICE_SOLD - Final_SalesXProduction_Month$REGRESSION_AVG_CPU

#add Revenue
Final_SalesXProduction_Month$MIN_TOTAL_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$MIN_PRICE_SOLD
Final_SalesXProduction_Month$AVG_TOTAL_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$AVG_PRICE_SOLD
Final_SalesXProduction_Month$MAX_TOTAL_REVENUE <- Final_SalesXProduction_Month$TOTAL_SHIPPED * Final_SalesXProduction_Month$MAX_PRICE_SOLD

#add profit
Final_SalesXProduction_Month$MIN_TOTAL_PROFIT <- Final_SalesXProduction_Month$MIN_TOTAL_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$AVG_TOTAL_PROFIT <- Final_SalesXProduction_Month$AVG_TOTAL_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$MAX_TOTAL_PROFIT <- Final_SalesXProduction_Month$MAX_TOTAL_REVENUE - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
Final_SalesXProduction_Month$ACTUAL_TOTAL_PROFIT <- Final_SalesXProduction_Month$TOTAL_SALES - Final_SalesXProduction_Month$ACTUAL_MONTHLY_COST
```

## ----------Export the Datasets for Visualizing-----------

```{r}
# write_csv(Final_SalesXProduction_Productline, file.choose(new = TRUE))
# write_csv(Final_SalesXProduction_Month, file.choose(new = TRUE))
```

## ---------- Bea's Production ---------------------

Forcast expected productline sales from june 2005 - December 2005

```{r}

# Group the data by date and product for time series
monthly_data <- sales_data %>%
  group_by(MONTH_ORDERED, PRODUCTLINE) %>%
  summarise(TOTAL_UNITS = sum(QUANTITYORDERED),
            .groups = "drop") %>%
   arrange(PRODUCTLINE, MONTH_ORDERED)

monthly_data$MONTH_ORDERED <- ym(monthly_data$MONTH_ORDERED)
  
#--------------Determine Which method to use per production line------------
production_model <- function(product_line) {
  # Create a productline time table
  submodel <- monthly_data %>%
    filter(PRODUCTLINE == product_line) %>%
    arrange(MONTH_ORDERED)
  
  # Build Time Series object
  ts_submodel <-ts(submodel$TOTAL_UNITS,
                   start = c(year(min(submodel$MONTH_ORDERED)),
                             month(min(submodel$MONTH_ORDERED))),
                   frequency = 12)
  last_date <- max(submodel$MONTH_ORDERED)
  target_end <- as.Date("2005-12-31")
  h <- interval(last_date, target_end) %/% months(1)
  
  mean_model <- meanf(ts_submodel, h = h)
  snaive_model <- snaive(ts_submodel, h = h)
  rwf_model <- rwf(ts_submodel, h = h, drift = TRUE)
  
  ets_ann <- ets(ts_submodel, model = "ANN")
  forecast_ets_ann <- forecast(ets_ann, h = h)
  
  ets_aaa <- ets(ts_submodel, model = "AAA")
  forecast_ets_aaa <- forecast(ets_aaa, h = h)
  
  ets_auto <- ets(ts_submodel)
  forecast_ets_auto <- forecast(ets_auto, h = h)
  
  arima_auto <- auto.arima(ts_submodel)
  forecast_arima <- forecast(arima_auto, h = h)
  
  lm_model <- tslm(ts_submodel ~ trend + season)
  forecast_lm <- forecast(lm_model, h = h)
  
  # STL needs 2 full years - safely skip if not enough data
  stl_forecast <- tryCatch(
    stlf(ts_submodel, h = h),
    error = function(e) NULL
  )
  
  accuracy_table <- list(
    Mean = accuracy(mean_model)[1,],
    Seasonal_Naive = accuracy(snaive_model)[1,],
    Random_Walk_Drift = accuracy(rwf_model)[1,],
    ETS_ANN = accuracy(forecast_ets_ann)[1,],
    ETS_AAA = accuracy(forecast_ets_aaa)[1,],
    ETS_Auto = accuracy(forecast_ets_auto)[1,],
    ARIMA = accuracy(forecast_arima)[1,],
    Linear_Regression = accuracy(forecast_lm)[1,]
  )
  if (!is.null(stl_forecast)) {
    accuracy_table$STL <- accuracy(stl_forecast)[1,]
  }
  
  accuracy_table <- round(do.call(rbind, accuracy_table), 2)
  cat("\n===", product_line, "===\n")
  print(accuracy_table)
  
}

#Run all forecasts
product_lines <- unique(monthly_data$PRODUCTLINE)
results <- map(product_lines, production_model)
results <- results[!sapply(results, is.null)]
```

#### Orders per product line Time series

```{r}
forecast_production <- function(product_line) {
  submodel <- monthly_data %>%
    filter(PRODUCTLINE == product_line) %>%
    arrange(MONTH_ORDERED) %>%
    # Fill in missing order points with 0's
    tidyr::complete(
      MONTH_ORDERED = seq.Date(min(MONTH_ORDERED), max(MONTH_ORDERED), by = "month"),
      fill = list(TOTAL_UNITS = 0)
    )
  
  ts_submodel <- ts(submodel$TOTAL_UNITS,
                 start = c(year(min(submodel$MONTH_ORDERED)),
                           month(min(submodel$MONTH_ORDERED))),
                 frequency = 12)
  # Calculate Time series ranges
  last_date <- max(submodel$MONTH_ORDERED)
  target_end <- as.Date("2005-12-31")
  h <- interval(last_date, target_end) %/% months(1)
  
  # Use different series depending num months in dataset
  if (length(ts_submodel) < 24){
    ts_forecast <- forecast(tslm(ts_submodel ~ trend + season), h = h) 
  } else{
    stl_model <- stl(ts_submodel, s.window = "periodic")
    ts_forecast <- forecast(stl_model, h=h)
  }
  # Display forecasting data
  cat("\n===", product_line, "===\n")
  print(ts_forecast)
  
  #plot the forecast
  autoplot(ts_forecast) +
    ggtitle(paste(product_line, "Orders Forecast"))
  
}

#Run all forecasts
product_lines <- unique(monthly_data$PRODUCTLINE)
#Map the products to the function
results <- map(product_lines, forecast_production)
# Omit null values
results <- results[!sapply(results, is.null)]
print(results)
```

## ---------- Elton's Production ---------------------

### EDA

#### Production Trends Over Time

```{r Load and Sort Data}
# Loading the data
df_total <- read_csv("monthly_salesXproduction.csv", show_col_types = FALSE)
df_line <- read_csv("monthly_salesXproduction_by_productline.csv", show_col_types = FALSE)

# Sort chronologically by month
df_total <- df_total %>% arrange(MONTH_PRODUCED)
df_line <- df_line %>% arrange(MONTH_PRODUCED)
```

```{r Total Monthly Production Over Time}
plot_total <- ggplot(df_total, aes(x = MONTH_PRODUCED, y = MONTHLY_PRODUCTION)) +
  geom_line(color = "blue", group = 1) +
  geom_point(color = "blue") +
  labs(title = "Total Monthly Production Over Time",
       x = "Month Produced",
       y = "Total Units Produced") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # Making x-axis labels more readable
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), 
                      by = max(1, floor(length(x)/15)))])
print(plot_total)
```

```{r Monthly Production by Product Line Over Time, echo = FALSE}
df_line_filtered <- df_line %>% 
  drop_na(factory_id) %>% 
  mutate(factory_label = paste("Factory", factory_id),
         MONTH_DATE = ym(MONTH_PRODUCED))

plot_productline <- ggplot(df_line_filtered, aes(x = MONTH_DATE, y = TOTAL_UNITS_PRODUCED,
                                        color = PRODUCTLINE, group = PRODUCTLINE)) +
  geom_line() +
  geom_point() +
  labs(title = "Monthly Production by Product Line Over Time",
       x = "Month Produced",
       y = "Units Produced",
       color = "Product Line") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_breaks = "2 months", date_labels = "%b `%y")

print(plot_productline)
```

```{r Monthly Production by Product Line Over Time (Faceted by Factory), echo = FALSE}
# Making cleaner labels
df_line_filtered <- df_line %>% 
  # Removing the factory with no data
  drop_na(factory_id) %>% 
  mutate(factory_label = paste("Factory", factory_id),
         MONTH_DATE = ym(MONTH_PRODUCED))

plot_productline_faceted <- ggplot(df_line_filtered, aes(x = MONTH_DATE, y = TOTAL_UNITS_PRODUCED,
                                        color = PRODUCTLINE, group = PRODUCTLINE)) +
  geom_line() +
  geom_point() +
  # Splitting the plot by factory
  facet_wrap(~ factory_label) +
  labs(title = "Monthly Production by Product Line Over Time (By Factory)",
       x = "Month Produced",
       y = "Units Produced",
       color = "Product Line") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        strip.text = element_text(face = "bold")) +
  # Easier to read label formatting, every four months for space savings
  scale_x_date(date_breaks = "4 months", date_labels = "%b `%y")

print(plot_productline_faceted)
```

#### Seasonality / Trend Patterns

```{r Time Series Decomposition for Seasonal Patterns and Trends}
# Making a column of Date objects
df_total = df_total %>% 
  mutate(MONTH_DATE = ym(MONTH_PRODUCED))

# Converting to a Time Series object
start_year <- year(min(df_total$MONTH_DATE))
start_month <- month(min(df_total$MONTH_DATE))

# Frequency of 12 for 12 months
production_ts <- ts(df_total$MONTHLY_PRODUCTION,
                    start = c(start_year, start_month),
                    frequency = 12)

# STL to split into Seasonality, Trend, and Remainder
decomposition <- stl(production_ts, s.window = "periodic")
plot_decomp <- autoplot(decomposition) +
  theme_minimal() +
  labs(title = "Time Series Decomposition of Total Monthly Production")

print(plot_decomp)
```

```{r Time Series Decomposition by Product Line}
# Ignoring product lines with no production, i.e. Trains
valid_product_lines <- df_line %>% 
  group_by(PRODUCTLINE) %>% 
  summarize(total = sum(TOTAL_UNITS_PRODUCED, na.rm = TRUE)) %>% 
  filter(total > 0) %>% 
  pull(PRODUCTLINE)

# Loop through each valid product line to generate plots
for (prod_line in valid_product_lines) {
  # Filter and prepare data for the specific product line
  df_sub <- df_line %>% 
    filter(PRODUCTLINE == prod_line) %>% 
    # If a month is missing, fill it with 0
    complete(MONTH_PRODUCED = unique(df_line$MONTH_PRODUCED),
             fill = list(TOTAL_UNITS_PRODUCED = 0)) %>% 
    arrange(MONTH_PRODUCED)
  
  # Create a Time Series object
  prod_ts <- ts(df_sub$TOTAL_UNITS_PRODUCED,
                start = c(start_year, start_month),
                frequency = 12)
  # Plot Decomposition
  decomp <- stl(prod_ts, s.window = "periodic")
  plot_decomp <- autoplot(decomp) +
    theme_minimal() +
    labs(title = paste("Time Series Decomposition of Monthly Production:", prod_line))
  
  print(plot_decomp)
  
  # Plot Seasonal Plot
  plot_season <- ggseasonplot(prod_ts, year.labels = TRUE, continuous = TRUE) +
  theme_minimal() +
  labs(title = paste("Seasonal Plot:", prod_line),
       y = "Units Produced",
       x = "Month")
  
  print(plot_season)
}
```

#### Descriptive Statistics

-   Factory

```{r}

```

-   Product Line

-   Month

```{r Monthly Production Statistics}
# Calculating average monthly production
cat("Average Monthly Production:\t", round(mean(df_total$MONTHLY_PRODUCTION), 2), "units\n")

# Find peak production month
peak_month <- df_total %>% filter(MONTHLY_PRODUCTION == max(MONTHLY_PRODUCTION))
cat("Peak Production Month:\t\t", peak_month$MONTH_PRODUCED, "with", peak_month$MONTHLY_PRODUCTION, "units\n")

# Find lowest production month
lowest_month <- df_total %>% filter(MONTHLY_PRODUCTION == min(MONTHLY_PRODUCTION))
cat("Lowest Production Month:\t", lowest_month$MONTH_PRODUCED, "with", lowest_month$MONTHLY_PRODUCTION, "units\n")

cat("Total Production by Product Line:\n")
product_line_summary <- df_line %>% 
  group_by(PRODUCTLINE) %>% 
  summarize(TOTAL_PRODUCED = sum(TOTAL_UNITS_PRODUCED, na.rm = TRUE)) %>% 
  arrange(desc(TOTAL_PRODUCED))

print(product_line_summary)
```

#### Discussion

-   Are trends visible?

    -   Yes, overall there is a strong upward trend for the first year and four months, followed by a leveling out and then small decrease.

-   Are there seasonal effects?

    -   All of the product lines show perfectly repeating seasonal patterns.

-   Do product lines differ substantially?

    -   The biggest outliers here are Ships with significantly higher production levels, and Classic Cars with significantly lower production levels. This is opposite of their relationship within sales data, so it is surprising.

## ---------- Kainen's Production ---------------------
